\begin{prob}
One of the interesting properties we had with functions in real
analysis was the notion of continuity. Is there a way to generalize
this notion to a topological setting?
\end{prob}
\begin{defn}{(Real Analysis Definition of Continuous)}
Let $f:X\to Y$ be a function, $X$ and $Y$ be sets. We say that $f$ is
``\define{Continuous}'' if for each $\varepsilon>0$ there is a
corresponding $\delta>0$ such that
\begin{equation}
|x-x_{0}|<\delta\Rightarrow|f(x)-f(x_0)|<\varepsilon
\end{equation}
Or in other words, for each $\varepsilon$ neighborhood of $f(x_0)$,
there is a $\delta$ neighborhood of $x_0$.
\end{defn}
\begin{rmk}
Observe that what we are doing is specifying a neighborhood in the
range of a certain size, then demanding there is a corresponding
neighborhood in the domain of a certain size. We can do this in real
analysis since the reals are sufficiently nice (they are a metric
space which is a really really strong condition for a topological
space). We want to generalize this for topological spaces so when we
work with metric spaces we recover our previous notion of continuity.
\end{rmk}
\begin{defn}
Let $X$, $Y$ be topological spaces, $f:X\to Y$ be a function. We say
that ``\define{$f$ is continuous at $x_0$}'' if for every neighborhood
$V$ of $f(x)$, there is a neighborhood $U$ of $x$ such that
$f(U)\subset V$.
\end{defn}
\begin{rmk}
This notion of continuity is nearly identical to the notion we
previously introduced from real analysis. The difference is that we
are not using a metric to keep track of every open neighborhood in the
domain and the range.

But observe we demand each open neighborhood $V$ of $f(x)$ has a
corresponding neighborhood $U$ in the domain such that the image of
$U$ is contained in $V$. This is precisely what we did in the real
analysis case, we specified an $\varepsilon$ neighborhood, and
demanded that for each $\varepsilon$ neighborhood there is a
corrsponding $\delta$ neighborhood in the domain such that the image
of the $\delta$ neighborhood is contained in the $\varepsilon$
neighborhood.
\end{rmk}

\begin{thm}
Let $X$ and $Y$ be topological spaces, if $f:X\to Y$ is such
that for each open set $V\subset Y$ its preimage $f^{-1}(V)\subset X$
is open, then $f$ is continuous.
\end{thm}
\begin{proof}
Trivial.
\end{proof}
\begin{rmk}{(Inverse Functions)}
Note that contrary to appearances \emph{continuity does not demand the
  function be invertible!} The preimage of a set \emph{is not} the
same as the function's inverse. Consider $f:X\to Y$, and $Z\subset Y$
is some open set. Then
\begin{equation}
f^{-1}(Z) = \left\{x\in X:\; f(x)\in Z\right\}
\end{equation}
which does not say anything about the existence of an inverse for $f$.
\end{rmk}

\begin{defn}
Let $X$, $Z$ be topological spaces, and $Y\subset X$ be a subspace. A
map
\begin{equation}
\begin{array}{llll}
i:& Y & \hookrightarrow& X\\
  & x & \mapsto& i(x)=x
\end{array}
\end{equation}
is defined as the ``\define{Inclusion Map}''. Similarly, a continuous map
\begin{equation}
\begin{array}{llll}
r:& X & \hookrightarrow& Y\\
  & x & \mapsto& r(x)=x\;\text{if }x\in Y
\end{array}
\end{equation}
is defined as the ``\define{Retraction}'' if the retraction of $i$ to
$Y$ is the identity on $A$.
\end{defn}
\begin{rmk}
We use the inclusion map and the retraction to extend and restrict
functions (respectively) by composing them with the functions of
interest. Note that the restriction of a continuous function
(i.e. composing it with a retraction) is the composition of two
continuous functions and thus continuous.
\end{rmk}
\begin{defn}
Let $X$, $Z$ be topological spaces, $Y\subset X$ be a subspace.
If $f:X\to Z$ is a function, we can define the ``\define{Restriction
  of $f$ to $Y$}'' as
\begin{equation}
f\circ r:X\to Z
\end{equation}
the composition of the retraction map (which is just
$\operatorname{id_Y}$ on $Y$) with $f$. 
\end{defn}
\begin{rmk}
Observe that the restriction of a function is just composition of
functions.
\end{rmk}
\begin{lem}{(Pasting Lemma)}
Let $X$ and $Y$ be topological spaces, and $U,$ $V$ be open sets in
$X$ such that $X=U\cup V$. Suppose $f:U\to Y$ and $g:V\to Y$ are
continuous and 
\begin{equation}
f(x)=g(x)\quad\text{for all }x\in U\cap V
\end{equation}
Then the function $h:X\to Y$ defined by
\begin{equation}
h(x) = \begin{cases} f(x),&\text{if }x\in U\\
g(x),&\text{if }x\in V
\end{cases}
\end{equation}
is continuous.
\end{lem}
