\chapter{Spin Groups and their Representations}

\begin{ddanger}
I have goofed up my conventions here, mixing up the orthonormal basis
$e_{i}$ for $V$ with the generators $\Gamma_{i}$ for $\Cl(V)$. When
revisiting my notes, I should fix this when possible. The reader should
understand this is what has happened, and your mileage may vary.
\end{ddanger}

\N{TODO: fix notation}
I should use $e_{\alpha}$ for an orthonormal basis, $\gamma_{\alpha}$
for the generators of $\Cl(V)$, and $\Gamma_{\alpha}$ for matrices in
the representation $\End(W)$ of $\Cl(p,q)$.

\N{TODO: super tensor products}
I believe the calculations below of the Clifford algebra irreducible
representations are correct, but I should be using the super tensor
product for the isomorphisms like
$\Cl(n+2)\iso\Cl(n)\super\otimes\Cl(2)$, and so on.

\section{Review of Clifford Algebras}

\M
Clifford algebras may be formed from a vector space $V$ over $\FF$. It's
especially intuitive when $V$ is finite-dimensional. We just take some
canonical basis $\Gamma_{1}$, \dots, $\Gamma_{n}$ for $V$, then we turn it into an
algebra by defining multiplication on vectors by linearity and the
following anticommutator relation:
\begin{equation}
\Gamma_{i}\Gamma_{j} + \Gamma_{j}\Gamma_{i} = -\delta_{i,j}.
\end{equation}
Observe that $\Gamma_{i}^{2}=-1$ and $\Gamma_{i}\Gamma_{j}=-\Gamma_{j}\Gamma_{i}$ when $i\neq j$.
The resulting algebra is precisely the \define{Clifford Algebra} $\Cl(V)$.

We can generalize this by altering the anticommutation relation to be
\begin{equation}
\Gamma_{i}\Gamma_{j} + \Gamma_{j}\Gamma_{i} = \eta_{i,j}
\end{equation}
for some matrix $\eta$. When $\eta_{i,j}=0$, we obtain the
\define{Grassmann Algebra}. We can also have ``split'' Clifford algebras
by having the signature of $\eta$ being indefinite.

\begin{remark}
Historically, people describe a Clifford algebra using a quadratic form
(a generalized ``norm squared'' quantity). This is completely equivalent
to what we have done. Namely, $\eta_{i,j}=-Q(\Gamma_{i})\delta_{i,j}$ where
$Q$ is the quadratic form on $V$.
\end{remark}

\N{Even Part}
The even part of the Clifford algebra is written as $\Cl^{0}(V)$ or
$\Cl(V)_{0}$. It consists of guys that look like
\[ a_{0}\cdot\mathbf{1} + a^{\mu\nu}\Gamma_{\mu}\Gamma_{\nu} + a^{\alpha\beta\mu\nu}\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\mu}\Gamma_{\nu}+\dots.\]
Each term has an even number of Clifford generators appearing in it. The
coefficients $a^{\mu\nu}$, $a^{\alpha\beta\mu\nu}$, etc., are completely
antisymmetric in all its indices; equivalently, we could simply sum over
$\alpha<\beta<\mu<\nu<\dots$. 

\begin{definition}
Let $A$ be an $\FF$-algebra (usually $\FF=\RR$, $\CC$, or $\HH$). An
\define{$A$-Representation} of the Clifford algebra $\Cl(V)$ is an algebra morphism
\[ \rho\colon\Cl(V)\to\hom_{A}(W, W) \]
where $W$ is a $A$-algebra representation, $\hom_{A}(W, W)$ is the space
of endomorphisms of $W$ commuting with the $A$-action.
\end{definition}

\N{Semisimple Algebras}
We call an algebra $A$ \define{Semisimple} if all its modules are
completely reducible (= direct sums of simple modules). This is one
possible criterion. We can consider $A$ as a left $A$-module when $A$ is
associative\footnote{If I recall correctly, we need associativity.}, and
then this coincides with $A$ as a semisimple $A$-module. Then $A$ is
semisimple if and only if the left regular representation for $A$ is
completely reducible.

In particular, $\Cl(V)$ is a semisimple algebra. Recall semisimple rings
generalize the concept of a group algebra $\CC[G]$ for finite groups
$G$. For a review of the structure theorem for Clifford algebras, see
\arXiv{1610.02418}. In particular, all irreducible representations of a
semisimple associative algebra may be found as subrepresentations of its
regular representation.\footnote{See Theorem 3.2.13 in \url{https://studenttheses.uu.nl/bitstream/handle/20.500.12932/29895/thesis.pdf?sequence=2}}

For more on this, see chapter 5 of Varadarajan~\cite{Varadarajan:2004yz}.
Specifically, when $V$ is even-dimensional, $\Cl(V)$ is semisimple ---
this is proven in Theorem~5.2.7, and the odd-dimensional case is proven
in Theorem~5.2.10.

\N{Notation: ``Volume Element''}
We write $\omega=\prod^{m}_{j=1}\Gamma_{j}\in\Cl(V)$ and refer to it as the
``\emph{Volume Element}''. Adams does not use this notation (or
terminology), but it caught on in the mathematical literature.
Physicists call it the \emph{Chiral Element}.

\N{Notation}
We write $\FF(n)$ for the algebra of $n\times n$ matrices with entries
in the division algebra $\FF$ (usually $\RR$, $\CC$, $\HH$).

We also write $\Cl_{p,q}$ for $\Cl(\RR^{p,q})$ using the indefinite norm
of signature $(p,q)$ with quadratic form
\begin{equation}
Q(x) = (x_{1})^{2}+\cdots+(x_{p})^{2}-(x_{p+1})^{2}-(\cdots)-(x_{p+q})^{2}.
\end{equation}
We write $\Cl_{p}=\Cl_{p,0}$. This is all fairly standard notation.

\begin{lemma}
Let $\FF=\RR$, $\CC$, or $\HH$.
Consider $\FF(n)$ as an $\RR$-algebra.
Then $\FF(n)$ has a natural representation on $\FF^{n}$ and, up to
equivalence, this is the only irreducible representation.
\end{lemma}

\begin{proof}
Matrix algebras are simple. Simple algebras have a single representation
up to isomorphism.
\end{proof}

\begin{proposition}
\begin{enumerate}
\item $\RR(n)\otimes\RR(m)\iso\RR(nm)$ for all $n$, $m\in\NN$
\item Extension of scalars: Let $\FF$ be $\CC$ or $\HH$ (or $\RR$). Then
  $\RR(n)\otimes_{\RR}\FF\iso\FF(n)$ for all $n\in\NN$
\item $\CC\otimes_{\RR}\CC\iso\CC\oplus\CC$
\item $\CC\otimes_{\RR}\HH\iso\CC(2)$
\item $\HH\otimes_{\RR}\HH\iso\RR(4)$.
\end{enumerate}
\end{proposition}

\begin{proof}
  (1) Obvious.

  (2) Also obvious, we just ``componentwise'' extend the scalars.

  (3) Write $\CC\iso\RR[x]/(x^{2}+1)$, then by direct calculation
\begin{calculation}
\CC\otimes_{\RR}\CC
\step[\iso]{since $\CC\iso\RR[x]/(x^{2}+1)$}
\CC\otimes_{\RR}\RR[x]/(x^{2}+1)
\step[\iso]{extension of scalars}
\CC[x]/(x^{2}+1)
\step[\iso]{Chinese remainder theorem}
\bigl(\CC[x]/(x+\I)\bigr)\oplus \bigl(\CC[x]/(x-\I)\bigr)
\step[\iso]{since $\CC[x]/(x\pm\I)\iso\CC$}
\CC\oplus\CC.
\end{calculation}

(4) Recall that $\HH$ is generated by $\I$ and $j$, since
  $k=ij$. Write
  \begin{equation}
\widehat{I} = \begin{bmatrix}\sqrt{-1} & 0\\
0 & \sqrt{-1}
\end{bmatrix},\quad\mbox{and}\quad
\widehat{J} = \begin{bmatrix} 0 & -1\\
1 & 0
\end{bmatrix}.
  \end{equation}
  Then we can observe $\widehat{I}\widehat{J}=-\widehat{J}\widehat{I}$
  squares to $-1$. Then
\begin{equation}
\begin{split}
  \HH\otimes_{\RR}\CC&\to\CC(2)\\
  \I\otimes1 &\mapsto\widehat{I}\\
  j\otimes1 &\mapsto\widehat{J}.
\end{split}
\end{equation}
This is a homomorphism and moreover an isomorphism.

(5) We have
\begin{equation}
\HH\otimes_{\RR}\HH\iso\End(\HH),
\end{equation}
and since $\HH\iso\RR^{4}$ as a vector space, this implies the result.
\end{proof}

\N{Facts}
We can compute:
\begin{subequations}
\begin{align}
\Cl_{1,0} &= \CC\\
\Cl_{0,1} &= \RR\oplus\RR\\
\Cl_{2,0} &= \HH\\
\Cl_{0,2} &= \RR(2)\\
\Cl_{1,1} &= \RR(2).
\end{align}
\end{subequations}
More helpfully, arranged as a table
\begin{equation}
\begin{array}{c|ccc}
(\Cl_{0,n})   & \huh   & \huh & \huh\\
  \RR(2)     & \huh   & \huh & \huh\\
\RR\oplus\RR & \RR(2) & \huh & \huh\\\hline
\RR          &  \CC   &  \HH & (\Cl_{m,0})
\end{array}
\end{equation}

\begin{theorem}
For all $n$, $r$, $s\geq0$, we have the following isomorphisms:
\begin{enumerate}
\item $\Cl_{n,0}\otimes\Cl_{0,2}\iso\Cl_{0,n+2}$
\item $\Cl_{0,n}\otimes\Cl_{2,0}\iso\Cl_{n+2,0}$
\item $\Cl_{r,s}\otimes\Cl_{1,1}\iso\Cl_{r+1,s+1}$.
\end{enumerate}
\end{theorem}

\begin{remark}
The third condition, $\Cl_{r,s}\otimes\Cl_{1,1}\iso\Cl_{r+1,s+1}$,
allows us to ``propagate'' a known entry ``diagonally'' across the table
of Clifford algebras.
\end{remark}

\N{Bott Periodicity}
We also have the following isomorphisms:
\begin{enumerate}
\item $\Cl_{n+8,0}\iso\Cl_{n,0}\otimes\Cl_{8,0}$
\item $\Cl_{0,n+8}\iso\Cl_{0,n}\otimes\Cl_{0,8}$
\item For complex Clifford Algebras: $\CCl_{n+2}\iso\CCl_{n}\otimes_{\CC}\CCl_{2}$.
\end{enumerate}

\M
This gives us the following table for real $\Cl_{p,q}$:
\begin{equation*}
  \begin{footnotesize}\hskip-3em
\begin{array}{c|ccccccccc}
  8 & \RR(16)            & \CC(16) & \HH(16) & \HH(16)\oplus\HH(16) & \HH(32) & \CC(64) & \RR(128) & \RR(128)\oplus\RR(128) & \RR(256)\\
  7 & \CC(8)             & \HH(8) & \HH(8)\oplus\HH(8) & \HH(16) & \CC(32) & \RR(64) & \RR(64)\oplus\RR(64) & \RR(128) & \CC(128)\\
  6 & \HH(4)             & \HH(4)\oplus\HH(4) & \HH(8) & \CC(16) & \RR(32) & \RR(32)\oplus\RR(32) & \RR(64) & \CC(64) & \HH(64)\\
  5 & \HH(2)\oplus\HH(2) & \HH(4) & \CC(8) & \RR(16) & \RR(16)\oplus\RR(16) & \RR(32) & \CC(32) & \HH(32) & \HH(32)\oplus\HH(32)\\
  4 & \HH(2)             & \CC(4) & \RR(8) & \RR(8)\oplus\RR(8) & \RR(16) & \CC(16) & \HH(16) & \HH(16)\oplus\HH(16)  & \HH(32)\\
  3 & \CC(2)             & \RR(4) & \RR(4)\oplus\RR(4) & \RR(8) & \CC(8) & \HH(8) & \HH(8)\oplus\HH(8) & \HH(16) & \CC(32)\\
  2 & \RR(2)             & \RR(2)\oplus\RR(2) & \RR(4) & \CC(4) & \HH(4) & \HH(4)\oplus\HH(4) & \HH(8) & \CC(16) & \RR(32)\\
  1 & \RR\oplus\RR       & \RR(2) & \CC(2) & \HH(2) & \HH(2)\oplus\HH(2) & \HH(4) & \CC(8) & \RR(16)  & \RR(16)\oplus\RR(16) \\
  0 & \RR                & \CC    & \HH & \HH\oplus\HH & \HH(2) & \CC(4) & \RR(8) & \RR(8)\oplus\RR(8) & \RR(16) \\\hline
    & 0                  & 1  & 2 & 3 & 4 & 5 & 6 & 7 & 8
\end{array}
  \end{footnotesize}
\end{equation*}

\begin{theorem}\label{thm:lie:spin:even-clifford-isomorphisms}
  \begin{enumerate}
  \item $\Cl_{n+1}^{0}\iso\Cl_{n}$ for each $n\in\NN_{0}$.
  \item $\Cl_{s,t}^{0}\iso\Cl_{s-1,t}$ for $s, t\in\NN$.
  \item $\Cl_{s,t}^{0}\iso\Cl_{t-1,s}$ for $s, t\in\NN$.
  \end{enumerate}
\end{theorem}

In particular, we have the following table describing the even
subalgebra as the following real associative algebras:

\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|} \hline
  $s-t$ mod $8$ & $\Cl_{s,t}^{0}$ & $N$ \\\hline
           1, 7 &       $\RR(N)$ & $2^{(s+t-1)/2}$\\
           3, 5 &       $\HH(N)$ & $2^{(s+t-3)/2}$\\ \hline
           2, 6 &       $\CC(N)$ & $2^{(s+t-2)/2}$\\
              4 & $\HH(N)\oplus\HH(N)$ & $2^{(s+t-4)/2}$\\
              0 & $\RR(N)\oplus\RR(N)$ & $2^{(s+t-2)/2}$\\ \hline
\end{tabular}
\caption{Even Clifford Subalgebras in $\RR^{s,t}$}
\end{table}

\section{The $\spin$ Representation}

\M
We consider the Clifford algebra $\Cl(\RR^{n})$ which has $n$ generators
$\Gamma_{1}$, \dots, $\Gamma_{n}$ which satisfy the usual relations
\begin{equation}\label{eq:spin:clifford-anticommutator}
\{\Gamma_{i}, \Gamma_{j}\} = \Gamma_{i}\Gamma_{j} + \Gamma_{j}\Gamma_{i} = -2\delta_{ij} = \eta_{ij}.
\end{equation}
If we wanted, we could replace the right-hand side by some nondegenerate
$n\times n$ matrix $\eta_{ij}$, but we will not need it for our purposes.
These generators $\Gamma_{i}$ generalize the Dirac gamma matrices. Then we
can consider the Lie algebra $\spin(n)$ spanned by $\Gamma_{i}\Gamma_{j}$ for
$i\neq j$.

\begin{lemma}\label{lemma:spin:clifford-algebra:commutator-of-one-and-two}
We have the commutator of generators for the Clifford algebra satisfy
$$[\Gamma_{\alpha}, \Gamma_{\beta}\Gamma_{\gamma}]=\Gamma_{\gamma}\eta_{\beta\alpha}-\Gamma_{\beta}\eta_{\alpha\gamma}.$$
\end{lemma}

\begin{proof}
By direct calculation,
\begin{calculation}
[\Gamma_{\alpha}, \Gamma_{\beta}\Gamma_{\gamma}]
\step{definition of commutator}
\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\gamma} - \Gamma_{\beta}\Gamma_{\gamma}\Gamma_{\alpha}
\step{associativity}
\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\gamma} - \Gamma_{\beta}(\Gamma_{\gamma}\Gamma_{\alpha})
\step{using anticommutation relations from Eq~\eqref{eq:spin:clifford-anticommutator}}
\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\gamma} - \Gamma_{\beta}(\eta_{\alpha\gamma}-\Gamma_{\alpha}\Gamma_{\gamma})
\step{distributivity}
\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\gamma} - \Gamma_{\beta}\eta_{\alpha\gamma}+\Gamma_{\beta}\Gamma_{\alpha}\Gamma_{\gamma}
\step{associativity}
\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\gamma} - \Gamma_{\beta}\eta_{\alpha\gamma}+(\Gamma_{\beta}\Gamma_{\alpha})\Gamma_{\gamma}
\step{using anticommutation relations from Eq~\eqref{eq:spin:clifford-anticommutator}}
\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\gamma} - \Gamma_{\beta}\eta_{\alpha\gamma}+(\eta_{\beta\alpha}-\Gamma_{\alpha}\Gamma_{\beta})\Gamma_{\gamma}
\step{distributivity}
\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\gamma} - \Gamma_{\beta}\eta_{\alpha\gamma}+\eta_{\beta\alpha}\Gamma_{\gamma}-\Gamma_{\alpha}\Gamma_{\beta}\Gamma_{\gamma}
\step{subtraction}
- \Gamma_{\beta}\eta_{\alpha\gamma}+\eta_{\beta\alpha}\Gamma_{\gamma}.\qedhere
\end{calculation}
\end{proof}

\begin{lemma}
The commutator obeys the product rule $[A,BC]=[A,B]C+B[A,C]$
and $[AB,C]=A[B,C] + [A,C]B$.
\end{lemma}

\begin{proposition}
We have the commutator of generators for the Clifford algebra satisfy
$$[\Gamma_{\alpha}\Gamma_{\beta}, \Gamma_{\gamma}\Gamma_{\delta}]=\bigl(\eta_{\beta\delta}\eta_{\gamma\alpha}-\eta_{\beta\gamma}\eta_{\alpha\delta}\bigr)
+\bigl(\Gamma_{\alpha}\Gamma_{\delta}\eta_{\gamma\beta}-\Gamma_{\alpha}\Gamma_{\gamma}\eta_{\beta\delta}
-\Gamma_{\beta}\Gamma_{\delta}\eta_{\gamma\alpha}
+\Gamma_{\beta}\Gamma_{\gamma}\eta_{\alpha\delta}\bigr).$$
\end{proposition}

\begin{proof}
By direct calculation,
\begin{calculation}
[\Gamma_{\alpha}\Gamma_{\beta}, \Gamma_{\gamma}\Gamma_{\delta}]
\step{using the product rule for commutator}
\Gamma_{\alpha}[\Gamma_{\beta}, \Gamma_{\gamma}\Gamma_{\delta}] + [\Gamma_{\alpha}, \Gamma_{\gamma}\Gamma_{\delta}]\Gamma_{\beta}
\step{using Lemma~\ref{lemma:spin:clifford-algebra:commutator-of-one-and-two}}
\Gamma_{\alpha}[\Gamma_{\beta}, \Gamma_{\gamma}\Gamma_{\delta}]
+ \bigl(\Gamma_{\delta}\eta_{\gamma\alpha}-\Gamma_{\gamma}\eta_{\alpha\delta}\bigr)\Gamma_{\beta}
\step{using Lemma~\ref{lemma:spin:clifford-algebra:commutator-of-one-and-two}}
\Gamma_{\alpha}\bigl(\Gamma_{\delta}\eta_{\gamma\beta}-\Gamma_{\gamma}\eta_{\beta\delta}\bigr)
+ \bigl(\Gamma_{\delta}\eta_{\gamma\alpha}-\Gamma_{\gamma}\eta_{\alpha\delta}\bigr)\Gamma_{\beta}
\step{distributivity}
\Gamma_{\alpha}\Gamma_{\delta}\eta_{\gamma\beta}-\Gamma_{\alpha}\Gamma_{\gamma}\eta_{\beta\delta}
+ \Gamma_{\delta}\Gamma_{\beta}\eta_{\gamma\alpha}-\Gamma_{\gamma}\Gamma_{\beta}\eta_{\alpha\delta}
\step{associativity}
\Gamma_{\alpha}\Gamma_{\delta}\eta_{\gamma\beta}-\Gamma_{\alpha}\Gamma_{\gamma}\eta_{\beta\delta}
+ (\Gamma_{\delta}\Gamma_{\beta})\eta_{\gamma\alpha}-(\Gamma_{\gamma}\Gamma_{\beta})\eta_{\alpha\delta}
\step{using anticommutation relations from Eq~\eqref{eq:spin:clifford-anticommutator}}
\Gamma_{\alpha}\Gamma_{\delta}\eta_{\gamma\beta}-\Gamma_{\alpha}\Gamma_{\gamma}\eta_{\beta\delta}
+ (\eta_{\beta\delta}-\Gamma_{\beta}\Gamma_{\delta})\eta_{\gamma\alpha}-(\Gamma_{\gamma}\Gamma_{\beta})\eta_{\alpha\delta}
\step{using anticommutation relations from Eq~\eqref{eq:spin:clifford-anticommutator}}
\Gamma_{\alpha}\Gamma_{\delta}\eta_{\gamma\beta}-\Gamma_{\alpha}\Gamma_{\gamma}\eta_{\beta\delta}
+ (\eta_{\beta\delta}-\Gamma_{\beta}\Gamma_{\delta})\eta_{\gamma\alpha}
-(\eta_{\beta\gamma}-\Gamma_{\beta}\Gamma_{\gamma})\eta_{\alpha\delta}
\step{distributivity}
\Gamma_{\alpha}\Gamma_{\delta}\eta_{\gamma\beta}-\Gamma_{\alpha}\Gamma_{\gamma}\eta_{\beta\delta}
+ \eta_{\beta\delta}\eta_{\gamma\alpha}-\Gamma_{\beta}\Gamma_{\delta}\eta_{\gamma\alpha}
-\eta_{\beta\gamma}\eta_{\alpha\delta}+\Gamma_{\beta}\Gamma_{\gamma}\eta_{\alpha\delta}
\step{collecting terms}
\bigl(\eta_{\beta\delta}\eta_{\gamma\alpha}-\eta_{\beta\gamma}\eta_{\alpha\delta}\bigr)
+\bigl(\Gamma_{\alpha}\Gamma_{\delta}\eta_{\gamma\beta}-\Gamma_{\alpha}\Gamma_{\gamma}\eta_{\beta\delta}
-\Gamma_{\beta}\Gamma_{\delta}\eta_{\gamma\alpha}
+\Gamma_{\beta}\Gamma_{\gamma}\eta_{\alpha\delta}\bigr).\qedhere
\end{calculation}
\end{proof}

\begin{remark}
We can interpret this as $\Gamma_{i}\Gamma_{j}$ satisfy the same commutation
relations as the generators $L_{ij}$ for the orthogonal Lie algebra.
\end{remark}

\M
What we tend to do is consider
\begin{equation}
\Gamma_{\alpha\beta} = \frac{1}{2}(\Gamma_{\alpha}\Gamma_{\beta} - \Gamma_{\beta}\Gamma_{\alpha}).
\end{equation}
Observe
\begin{equation}
\Gamma_{\alpha}\Gamma_{\beta} = \frac{1}{2}(\Gamma_{\alpha}\Gamma_{\beta} -
\Gamma_{\beta}\Gamma_{\alpha}) + \frac{1}{2}(\Gamma_{\alpha}\Gamma_{\beta} + \Gamma_{\beta}\Gamma_{\alpha})
= \Gamma_{\alpha\beta} - \eta_{\alpha\beta}\mathbf{1}.
\end{equation}
%% It is not hard to find
%% \begin{equation}
%% \begin{split}
%% \Gamma_{\alpha\beta} &= \frac{1}{2}(\Gamma_{\alpha}\Gamma_{\beta} + \Gamma_{\alpha}\Gamma_{\beta} - \eta_{\alpha\beta})\\
%% &= \Gamma_{\alpha}\Gamma_{\beta} - \frac{1}{2}\eta_{\alpha\beta}.
%% \end{split}
%% \end{equation}
Why do this?

Well, consider the representation of $\so(n)$ using 2-forms, writing
\begin{equation}
\rho(a) = a_{\mu\nu}\D x^{\mu}\wedge\D x^{\nu},
\end{equation}
using summation convention (repeated indices, with one downstairs
[subscript] and another upstairs [superscript], are summed over). Then
$a_{\mu\nu}$ is antisymmetric, which means it belongs to $\so(n)$. Thus
$\rho\colon\so(n)\to\Exterior^{2}\CC^{n}$ is a faithful representation.

But we're not done: identify $\D x^{\mu}\wedge\D x^{\nu}$ with
$\Gamma_{\mu\nu}$, and we obtain the $\spin(n)$ representation for $\so(n)$.
To be clear, we have a mapping
\begin{equation}
a_{\mu\nu}\D x^{\mu}\wedge\D x^{\nu}\mapsto\sum_{\mu,\nu}a_{\mu\nu}\Gamma_{\mu\nu}.
\end{equation}
This produces the spin representation
\begin{equation}
\rho(a) = \sum_{\mu,\nu}a_{\mu\nu}\Gamma_{\mu\nu}.
\end{equation}
It turns out to be quite useful.

\section{Spin Group}

\begin{definition}
Let $V$ be a finite-dimensional vector space over $\RR$ or $\CC$.
The \define{Pin Group} $\Pin(V)$ consists of elements of $\Cl(V)$
with unit norm, multiplication induced from $\Cl(V)$.
\end{definition}

\M
The inverse of $c\, \Gamma_{1}(\cdots)\Gamma_{k}\in\Pin(V)$ (for some nonzero $c\in\FF$)
is $(-1)^{k}c^{-1}\Gamma_{k}(\cdots)\Gamma_{2}\Gamma_{1}$ when we have $\Gamma_{i}^{2}=-1$
for each $i=1$, \dots, $n$.

Why would we have a coefficient $c\neq\pm1$ at all? Well, if we
consider, e.g.,
\begin{equation}
\begin{pmatrix}
v_{1}\\ v_{2}
\end{pmatrix} = \begin{pmatrix}\cos(\theta) & -\sin(\theta)\\
\sin(\theta) & \cos(\theta)
\end{pmatrix}
\begin{pmatrix}
\Gamma_{1}\\ \Gamma_{2}
\end{pmatrix},
\end{equation}
then $v_{1}$ is length 1 and inverible. But it is the sum of two
generators with nontrivial coefficients (for $\theta\notin\pi\QQ$, for
example). This is an example of a random element which is a pinor
(element of $\Pin(V)$).

\begin{remark}
Adams~\cite{adams1996:ex} gives an equivalent construction of the Pin
group, but is rather confusing: he uses $\beta$ to reverse the order of
multiplication, for example, instead of using the transpose of the
representation. Adams's $\alpha$ simply maps $\Gamma_{i}\mapsto-\Gamma_{i}$ on the
canonical generators for $\Cl(V)$. His
$\gamma=\alpha\circ\beta=\beta\circ\alpha$ called \define{Conjugation}
by DJH Garlin~\cite[pg.94]{Garling:2011zz}. Observe that $\alpha$ acts
trivially on even guys, and only on odd guys will something nontrivial
happen. 
\end{remark}

\begin{remark}
Adams defines $\Pin(V)$ as those $x\in\Cl(V)$ such that:
\begin{enumerate}
\item $x(\gamma(x))=(\gamma(x))x=1$, i.e., its conjugate is its inverse; and
\item the map $\pi_{x}\colon V\subset\Cl(V)\to\Cl(V)$ defined by $\pi_{x}(v)=xv(\beta(x))$
is injective.
\end{enumerate}
Then he proves $\pi$ is a 2-to-1 morphism of $\Pin(V)\onto\O(V)$. The
preimage of $\SO(V)$ is then defined as $\Spin(V)$.

This coincides with what Garlin~\cite[ch.8]{Garling:2011zz} does, since
$x^{-1}=\gamma(x)$ and so Garlin's approach [rewritten to use Adams's
  notation] using $v\mapsto x v \alpha(x^{-1})$, and direct calculation
shows
\begin{calculation}
  x v \alpha(x^{-1})
\step{since $x^{-1} = \gamma(x)$ by Adams's first criterion}
  x v ((\alpha\circ\gamma)(x))
\step{since $\gamma=\alpha\circ\beta$}
x v ((\alpha\circ\alpha\circ\beta)(x))
\step{since $\alpha\circ\alpha=\id$}
xv(\beta(x)).
\end{calculation}
But Garlin is slightly more general.
\end{remark}

\N{Adjoint Action of $\Pin(V)$ on $V$}
We can construct a representation of $\Pin(V)$ using $V$ by means of
\begin{equation}
\begin{split}
  \Ad\colon\Pin(V)&\to\aut(V)\\
  x&\mapsto\Ad_{x}
\end{split}
\end{equation}
where $\pi_{x}$ is defined as acting on $v\in V$ by using the inclusion
$i\colon V\into\Cl(V)$ as
\begin{equation}
\Ad_{x}(v) = x\,i(v)\,x^{-1}.
\end{equation}
If we observe that $x^{-1}=-x/Q(x)$ using the quadratic form $Q(-)$ [for
$x\in\Pin(V)$, we have $Q(x)=\pm1$],
then
\begin{calculation}
\Ad_{x}(v)
\step{using the definition of the adjoint representation}
xvx^{-1}
\step{since $x^{-1}=-x/Q(x)$}
\frac{-1}{Q(x)}xvx
\step{associativity}
\frac{-1}{Q(x)}(xv)x
\step{using anticommutator relations and induced bilinear form $B(-,-)$}
\frac{-1}{Q(x)}(-vx - 2B(v,x)\mathbf{1})x
\step{distributivity}
\frac{-1}{Q(x)}(vQ(x) - 2B(v,x)x)
\step{scalar distributivity}
-v + 2\frac{B(v,x)}{Q(x)}x
\end{calculation}
The reader may confirm this is, in fact, the negation of the reflection
of $v$ about $x$. We will write this as
\begin{equation}
\Ad_{x}(v) = -R_{x}(v)
\end{equation}
for $R_{x}$ be the linear operator encoding reflection about the
hyperplane perpendicular to $x$.

\N{Twisted Adjoint Representation}
We can construct the Twisted Adjoint action $\widetilde{\Ad}_{x}(v)=(-x)vx^{-1}$,
or more generally $\widetilde{\Ad}_{x}(v)=\alpha(x)\,v\,x^{-1}$ using
Adams's notation. This coincides with $\pi_{x}(v)$. When $a=u_{1}\dots u_{k}\in\Pin(V)$,
we have
\begin{equation}
\widetilde{\Ad}_{a}=R_{u_{1}}\circ(\cdots)\circ R_{u_{k}}.
\end{equation}
Reflections are orthogonal transformations, therefore $\widetilde{\Ad}$
defines a group morphism
\begin{equation}
\widetilde{\Ad}\colon\Pin(V)\to\O(V).
\end{equation}
We also see that $\widetilde{\Ad}$ is surjective by the Cartan--Dieudonn\'e
theorem.

\begin{theorem}[Cartan--Dieudonn\'e]
Every $g\in\O(V)$ is the product of a finite number of reflections
$R_{u_{1}}\circ(\cdots)\circ R_{u_{k}}$ along non-null lines (i.e.,
$Q(u_{i})\neq0$ for each $i=1,\dots,k$) and moreover $k\leq\dim(V)$.
\end{theorem}

\begin{theorem}
  The following is an exact sequence:
  \begin{equation}
1\to\{\pm1\}\into\Pin(V)\xonto{\widetilde{\Ad}}\O(V)\to1.
  \end{equation}
\end{theorem}

\N{Spin Group}
When we restrict attention to $\Pin(V)\cap\Cl^{0}(V)$ elements with an
even number of factors, then we obtain the spin group, denoted $\Spin(V)$.
This is a double cover of $\SO(V)$, everything works out nicely.

\N{Lie Algebra of $\Pin(V)$}
For $V\iso\FF^{n}$ (here $\FF$ being $\RR$ or $\CC$), $\Pin(V)$ has an
associated Lie algebra spanned by $\Gamma_{i}\Gamma_{j}\in\Cl(V)$ for $i<j$, and
whose Lie bracket is the commutator.

\M We can then consider the subgroup of $\Pin(V)$ consisting of an even
number of generators multiplied together. This is precisely the Spinor
group $\Spin(V)$.

Observe, this works out nicely, since the exponentiation of the Lie
algebra for $\Pin(V)$ should give us the component connected to the
identity, which is precisely $\Spin(V)$.

\M
Observe that, for $i < j$, we have:
\begin{calculation}
  (\Gamma_{i}\Gamma_{j})^{2}
\step{unfold}
  \Gamma_{i}\Gamma_{j}\Gamma_{i}\Gamma_{j}
\step{since $i\neq j$ we have $\Gamma_{j}\Gamma_{i}=-\Gamma_{i}\Gamma_{j}=(-1)\Gamma_{i}\Gamma_{j}$}
  \Gamma_{i}(-1)\Gamma_{i}\Gamma_{j}\Gamma_{j}
\step{using $\Gamma_{j}^{2}=-1$ and $\Gamma_{i}(-1)\Gamma_{i}=+1$}
  (+1)(-1) = -1.
\end{calculation}
Therefore for any $t\in\RR$, we have:
\begin{equation}
\E^{\Gamma_{i}\Gamma_{j}t} = \cos(t) + \Gamma_{i}\Gamma_{j}\sin(t),
\end{equation}
using Euler's formula. But when acting on $V$, we see that
$x=\exp(\Gamma_{i}\Gamma_{j}t)$ is represented by the matrix
\begin{equation}%\renewcommand{\arraystretch}{1.3}
\begin{split}
  \phantom{\pi_{x}} &\\
  \pi_{x}&=\quad \left[\begin{array}{ccccccccccc}
% 1 & 2    & 3 &  4       & 5 &  6     & 7 & 8         & 9 & 10 & 11
1 &        &   & \UP{i}{} &   &        &   & \UP{j}{}  &   &        & \\
  & \ddots &   &          &   &        &   &           &   &        & \\
  &        & 1 &          &   &        &   &           &   &        & \\
\LF{i}  &        &   & \cos(2t) &   &        &   & -\sin(2t) &   &        & \\
  &        &   &          & 1 &        &   &           &   &        & \\
  &        &   &          &   & \ddots &   &           &   &        & \\
  &        &   &          &   &        & 1 &           &   &        & \\
\LF{j}  &        &   & \sin(2t) &   &        &   & \cos(2t)  &   &        & \\
  &        &   &          &   &        &   &           & 1 &        & \\
  &        &   &          &   &        &   &           &   & \ddots & \\
  &        &   &          &   &        &   &           &   &        & 1
  \end{array}\right]
\end{split}
\end{equation}
Believe me? Well, even I have my doubts, so let us see how it acts on a
unit vector $\Gamma_{k}\in V$:
\begin{calculation}
(\cos(t) + \Gamma_{i}\Gamma_{j}\sin(t))\Gamma_{k}(\cos(t) - \Gamma_{i}\Gamma_{j}\sin(t))
\step{distributivity}
(\Gamma_{k}\cos(t) + \Gamma_{i}\Gamma_{j}\Gamma_{k}\sin(t))(\cos(t) - \Gamma_{i}\Gamma_{j}\sin(t))
\step{distributivity}
\Gamma_{k}\cos^{2}(t) + \Gamma_{i}\Gamma_{j}\Gamma_{k}\sin(t)\cos(t)
-\Gamma_{k}\Gamma_{i}\Gamma_{j}\cos(t)\sin(t) - \Gamma_{i}\Gamma_{j}\Gamma_{k}\Gamma_{i}\Gamma_{j}\sin^{2}(t)
\end{calculation}
Now, if $\Gamma_{k}=\Gamma_{i}$ or if $\Gamma_{k}=\Gamma_{j}$, then we have
\begin{calculation}
(\cos(t) + \Gamma_{i}\Gamma_{j}\sin(t))\Gamma_{k}(\cos(t) - \Gamma_{i}\Gamma_{j}\sin(t))
\step{continuing from last step}
\Gamma_{k}\cos^{2}(t) + \Gamma_{i}\Gamma_{j}\Gamma_{k}2\sin(t)\cos(t) - \Gamma_{i}\Gamma_{j}\Gamma_{k}\sin^{2}(t)
\step{using trig identities}
\Gamma_{k}\cos(2t) + \Gamma_{i}\Gamma_{j}\Gamma_{k}\sin(2t)
\step{collecting terms}
\bigl(\cos(2t) + \Gamma_{i}\Gamma_{j}\sin(2t)\bigr)\Gamma_{k}.
\end{calculation}
Hence we find if $\Gamma_{k}=\Gamma_{i}$ or if $\Gamma_{k}=\Gamma_{j}$,
\begin{subequations}
\begin{equation}
(\cos(t) + \Gamma_{i}\Gamma_{j}\sin(t))\Gamma_{k}(\cos(t) - \Gamma_{i}\Gamma_{j}\sin(t))
= \bigl(\cos(2t) + \Gamma_{i}\Gamma_{j}\sin(2t)\bigr)\Gamma_{k}.
\end{equation}
If $\Gamma_{k}\neq \Gamma_{i}$ and $\Gamma_{k}\neq \Gamma_{j}$, then we find by a
straightforward calculation (and using the $\cos^{2}(t)+\sin^{2}(t)=1$
trig identity):
\begin{equation}
(\cos(t) + \Gamma_{i}\Gamma_{j}\sin(t))\Gamma_{k}(\cos(t) - \Gamma_{i}\Gamma_{j}\sin(t))=\Gamma_{k}.
\end{equation}
\end{subequations}
Combining these two results, and rewriting it using a matrix, we obtain
precisely the desired matrix.

Coincidentally, this is why we have a 2-to-1 covering of $\SO(n)$
\textbf{as groups}, because we have $\cos(2t)$ and $\sin(2t)$ as the
entries of our matrix. But \textbf{as Lie algebras}, we have an
isomorphism $\spin(n)\iso\so(n)$.

\N{Maximal Torus of $\Spin(V)$}
The maximal torus $T$ for $\Spin(V)$, $V\iso\RR^{m}$, are elements of
the form
\begin{equation}
y = \prod^{n}_{r=1}[\cos(x_{r}/2) + \Gamma_{2r-1}\Gamma_{2r}\sin(x_{r}/2)]
\end{equation}
where $x_{r}\in\RR$ are parameters (``Euler angles''). These correspond
to matrices
\begin{equation}
  \pi_{y} = 
\begin{pmatrix}
\cos(x_{1}) & -\sin(x_{1}) &            &              & & & & \\
\sin(x_{1}) &  \cos(x_{1}) &            &              & & & & \\
           &              & \cos(x_{2}) & -\sin(x_{2}) & & & & \\
           &              & \sin(x_{2}) &  \cos(x_{2}) & & & & \\
 & & & & \ddots &            &              & \\
 & & & &        & \cos(x_{n}) & -\sin(x_{n}) & \\
 & & & &        & \sin(x_{n}) &  \cos(x_{n}) &  \\
 & & & &        &             &             & 1
\end{pmatrix}
\end{equation}
If $m=2n$, we just delete the last row and column (i.e., the row and
column with $1$ in the lower right corner).

\begin{remark}
Adams~\cite{adams1996:ex} concludes chapter 3 with a remark proving, for
$\RR^{n}$ guys, every element of $\Spin(\RR^{n})$ is conjugate to an
element in its torus. This seems intuitive, but I may be missing
something. 
\end{remark}

\section{Representations}

\M
Since $\Spin(V)\subset\Cl^{0}(V)$ is a subset of even Clifford
subalgebra, then any $\Cl^{0}(V)$-module is a representation of
$\Spin(V)$. In fact, irreducible representations of $\Cl_{r,s}^{0}$ 
induces irreducible representations for $\Spin_{r,s}$.

\begin{proposition}[{Adams~\cite[Prop.4.1]{adams1996:ex}}]
The algebras $\Cl(V)$ and $\Cl^{0}(V)$ are semi-simple, and hence all
their representations are completely reducible.
\end{proposition}

Adams proves the representations are completely reducible.

\begin{proof}[Proof sketch]
Adams begins by considering the standard basis $\Gamma_{1}$, \dots, $\Gamma_{m}$
for $V\iso\FF^{m}$, then constructs the subgroup
\begin{equation}
E = \{\pm\prod^{m}_{j=1}e^{i_{j}}_{j} \mid i_{j}=0\mbox{ or }1\},
\end{equation}
considered as a subgroup of order $2^{m+1}$ of $\Cl(V)$, which should
correspond to matrices $\diag(\pm1,\dots,\pm1)$ of $\O(m)$.

Then in $\Cl^{0}(V)$ we have the subgroup $E_{0}$ of $2^{m}$ elements
with $\sum_{j}i_{j}$ being even, specifically we have $E_{0}\subset E\subset\Pin(V)$.

We construct an operator $\nu$ corresponding to $-1\in\Cl^{0}(V)$
considered as an element of $E_{0}$ (or $E$ or $\Pin(V)$).

Adams observes $\nu$ is the generator of the $\pi$ representation's kernel.

Now, there are two claims made:
\begin{enumerate}
\item A module over $\Cl(V)$ gives a representation of $E$ where $\nu$
  acts as $-1$.
\item A representation of $E$ where $\nu$ acts as $-1$ gives a module
  over $\FF[E]/(\nu+1)\iso\Cl(V)$ the quotient of the group algebra
  $\FF[E]$ modulo $\nu+1\sim0$ (i.e., $\nu\sim-1$).
\end{enumerate}
In particular, these claims tell us the representation theory of
$\Cl(V)$ may be deduced from the representations of the finite group $E$.
This means all representations are completely reducible. A similar
argument holds for $\Cl^{0}(V)$ using $E_{0}$ instead of $E$.
\end{proof}

\begin{remark}
Adams will make heavy usage of these groups $E$, $E_{0}$, and the $\nu$ operator.
\end{remark}

\begin{remark}
Crucially, the insight that $\FF[E]/(\nu+1)\iso\Cl(V)$ allows us to find
representations for $E$, then induce a representation for $\Cl(V)$. This
is used in several proofs. The mnemonic appears to be that ``$\nu$ is
\emph{\textbf{n}egative one}.''
\end{remark}

\begin{lemma}\label{lemma:spin:center-of-e0}
The center of $E_{0}$ is $\{\pm1\}$ if $m=2n+1$ is odd dimensional, and
$\{\pm1,\pm\prod^{m}_{j=1}e_{j}\}$ if $m=2n$ is even dimensional.
\end{lemma}

\begin{proof}
Recall, the group center consists of $z\in Z(G)$ such that for all $g\in G$,
conjugation $z^{g}=g^{-1}zg=z$. It suffices to check conjugation by
every generator of the group $E_{0}$, which are precisely $\Gamma_{r}\Gamma_{s}$. Now if we conjugate
$g=\prod^{m}_{j=1}\Gamma_{j}^{i_{j}}$ by $\Gamma_{r}\Gamma_{s}$ where $i_{r}=1$ and $i_{s}=0$,
we change its sign. So if $g\in Z(E_{0})$, then $g=\pm1$ or $\pm \Gamma_{1}\Gamma_{2}\dots \Gamma_{m}$.
The latter is in the center only for $m$ even.
\end{proof}

\N{Fact}\label{fact:spin:complex-irrep-abelian-group-one-dim} Every $\CC$ irreducible representation for Abelian groups are
one-dimensional.

\N{Induced Representations}\label{chunk:spin:induced-rep}
Just to review induced representations, let $G$ be a finite group, let
$H\leq G$ be any subgroup, let $[G : H]=n$.
Suppose we have a representation $\pi\colon H\to\aut(V)$.
Suppose we have a complete set of coset representatives $g_{1}$, \dots,
$g_{n}$ for $H$ in $G$.

First, we should observe, for any $g\in G$ and any coset representative
$g_{i}$, we have $gg_{i}\in G$ (by closure of
multiplication). Therefore, $gg_{i}$ lives in a left coset of $H$, say
$g_{j}H$. More explicitly, $gg_{i}=g_{j}h_{i}$ for some $h_{i}\in H$.
This can be done for each coset representative, giving us a function of
indices $j(i)$, and we write:
\begin{equation}
g\cdot g_{i} = g_{j(i)}h_{i}.
\end{equation}
So far, so good.

Now we construct our representation space,
\begin{equation}
\widetilde{V} = \bigoplus^{n}_{i=1}g_{i}V.
\end{equation}
Then we define a $G$-action on $\widetilde{V}$ by using the insights
from the previous paragraph:
\begin{equation}
g\cdot\sum^{n}_{i=1}g_{i}\vec{v}_{i}=\sum^{n}_{i=1}g_{j(i)}\bigl(\pi(h_{i})\vec{v}_{i}\bigr).
\end{equation}
The resulting representation for $G$ is denoted $\operatorname{Ind}_H^G\pi$
or $\operatorname{Ind}_H^GV$ depending on author.

We can generalize this construction, using the insight that --- as a
representation space for $H$, $V$ is a $\FF[H]$-module --- the induced
representation is precisely the tensor product
\begin{equation}
\operatorname{Ind}_H^GV = \FF[G]\otimes_{\FF[H]}V.
\end{equation}
No conditions of finiteness are necessary in this generalized version.

\M\label{chunk:spin:induced-rep-as-hom-set} Briefly, we can see that $\operatorname{Ind}_H^GV$ is isomorphic to
the set
\begin{equation}
\hom_{H}(G,V) = \{f\colon G\to V\mid \forall h\in H,\forall g\in G,
f(hg)=h\cdot f(g)\}.
\end{equation}
The $G$-action on this set is
\begin{equation}
g'\cdot f(g) = f(gg').
\end{equation}
This is a more useful characterization for Adams's next proposition.

\N{Aside on Spin Representations}
Note that an irreducible representation for $\Cl^{0}(V)$ restrict to an
irreducible representation for $\Spin(V)$. It's for this reason that
Adams constructs an irrep of $\Cl^{0}(V)$ in the next proposition's
proof, and then sweeps the $\Spin(V)$ details under the rug.

\begin{proposition}[{Adams~\cite[Prop.4.2]{adams1996:ex}}]
\begin{enumerate}
\item If $m=\dim(V)=2n+1$ is odd, then $\Cl^{0}(V)$ has one irreducible
  representation $\Delta$ of degree $2^{n}$ affording a representation
  $\Delta$ of $\Spin(2n+1)$ with weights $\frac{1}{2}(\pm x_{1}\pm x_{2}\pm\cdots\pm x_{n})$
  with each sign being chosen independently, so there are $2^{n}$ of
  these weights.
\item If $m=\dim(V)=2n$ is even, then $\Cl^{0}(V)$ has two irreducible
  representations $\Delta^{+}$, $\Delta^{-}$ of degree $2^{n-1}$
  affording representations $\Delta^{+}$, $\Delta^{-}$ of $\Spin(2n)$
  with weights $\frac{1}{2}(\pm x_{1}\pm x_{2}\pm\cdots\pm x_{n})$
  with an even number of $-$ minus signs for $\Delta^{+}$ and an odd number of
  $-$ minus signs for $\Delta^{-}$. There are $2^{n-1}$ weights.
\item If $V$ is a complex vector space, then these constructions give
  complex-analytic representations for $\Spin_{\CC}(m)$.
\end{enumerate}
\end{proposition}

\begin{danger}
We will construct \emph{complex representations} for a \emph{real}
Clifford algebra which restricts to \emph{complex representations} for a
\emph{real} Spin group. This is extraordinarily confusing at first
glance, unless one is forewarned about this scheme.
\end{danger}

Physicists call $\Delta^{+}$ and $\Delta^{-}$ the \define{Half-Spin Representations},
some (e.g., Deligne) call them the \define{Semi-Spin Representations}.

\begin{remark}
It appears, thanks to Theorem~\ref{thm:lie:spin:even-clifford-isomorphisms},
an irreducible representation for $\Cl_{s,t}^{0}$ could be obtained from
$\Cl_{s-1,t}$. 
\end{remark}

\begin{remark}
I haven't seen a good explanation of Adams's proof, but it appears to be
fleshed out a bit more fully here: \url{https://www2.math.upenn.edu/~brweber/Courses/2013/Math651/Notes/L18_CliffSpin.pdf}
\end{remark}

\begin{remark}
This proposition is used later when constructing a Lie algebra of type
$\mathtt{E}_{8}$. 
\end{remark}

Adams's proof is rather convoluted, but amounts to constructing an
induced representation for the finite group $E_{0}$, then observes it
gives an irreducible representation for $\Spin(V)$.

\begin{proof}[Proof sketch]
By Schur's lemma, $\nu$ acts on any irreducible representation as either
$1$ or $-1$. The representations in which it acts as $1$ are
representations of $E_{0}/\langle\nu\rangle$ which is an Abelian group
of order $2^{m-1}$. Therefore there are $2^{m-1}$ one-dimensional
representations of $E_{0}$ in which $\nu$ acts as $1$.

Since the kernel of $E_{0}\to E_{0}/\langle\nu\rangle$ has exactly two
elements, the conjugacy classes in $E_{0}$ are either one element (if
the element is central) or two elements $\pm g$.

Recall: isomorphism classes of irreducible representations (over $\CC$)
of a finite group are in one-to-one correspondence [i.e., bijective with]
the conjugacy classes.

We see that $E_{0}$ has one (resp., two) more irreducible class(es) of
representation(s) than $E_{0}/\langle\nu\rangle$ if $m=2n+1$ (resp., $m=2n$).

Let $F\subset E_{0}$ be the subgroup generated by $\Gamma_{1}\Gamma_{2}$,
$e_{3}e_{4}$, \dots, $e_{2r-1}e_{2r}$, \dots, $e_{2n-1}e_{2n}$. This is
an Abelian group of order $2^{n+1}$, so
\begin{equation}
[E_{0} : F] = \begin{cases}2^{n} &\mbox{if }m=2n+1\\
2^{n-1} &\mbox{if }m=2n.
\end{cases}
\end{equation}
Is this really Abelian? Yes, we can calculate (assuming $i\neq j\neq k\neq\ell$):
\begin{calculation}
  e_{i}e_{j}e_{k}e_{\ell}
\step{anticommutator}
  -e_{i}e_{k}e_{j}e_{\ell}
\step{anticommutator}
  e_{i}e_{k}e_{\ell}e_{j}
\step{anticommutator}
  -e_{k}e_{i}e_{\ell}e_{j}
\step{anticommutator}
  e_{k}e_{\ell}e_{i}e_{j}.
\end{calculation}
Now we will begin to construct the induced representations.

We know (\S\ref{fact:spin:complex-irrep-abelian-group-one-dim})
every complex irreducible representation for $F$ will be one-dimensional.
Choose a complex one-dimensional representation $W$ of $F$ in which
$\nu$ acts as $-1$, and $e_{2r-1}e_{2r}$ acts as $\I\varepsilon_{r}$,
where $\varepsilon_{r}=\pm1$ and $\I=\sqrt{-1}$.

Form the induced representation (\S\ref{chunk:spin:induced-rep})
\begin{equation}
\operatorname{Ind}^{E_{0}}_{F}W = \CC[E_{0}]\otimes_{\CC[F]}W,
\end{equation}
where $\CC[E_{0}]$ is the complex group algebra for $E_{0}$.

Since $\deg(\operatorname{Ind}^{E_{0}}_{F}W)=[E_{0}:F]\deg(W)$ and
$\deg(W)=1$, we find the induced representation for $E_{0}$ is degree
$2^{n}$ for $m=2n+1$ and degree $2^{n-1}$ for $m=2n$. Observe, since
$\nu$ acts as $-1$ on $W$, that in the induced representation $\nu$
continues to act as $-1$.

Adams argues without proof that the basis for the representation is
given by
\begin{subequations}
\begin{align}
\prod^{2n+1}_{\substack{i=1\\i~\text{odd}}}e_{i}^{j_{i}}\quad\mbox{with}\quad\sum^{2n+1}_{1}j_{i}~\mbox{even},\quad\mbox{if }
m=2n+1,\\
\intertext{and}
\prod^{2n-1}_{\substack{i=1\\i~\text{odd}}}e_{i}^{j_{i}}\quad\mbox{with}\quad\sum^{2n-1}_{1}j_{i}~\mbox{even},\quad\mbox{if }
m=2n.
\end{align}
\end{subequations}
These are precisely the elements of $E_{0}$ which are not elements of $F$.

Now, there are two happy endings depending on $m$. The story is
basically the same: we will obtain an irreducible representation (or
two) for $E_{0}$ and this induces an irreducible representation for
$\Spin(V)$. And they live happily ever after.

\medbreak
\textbf{Case 1: $m=2n+1$}. There are $2^{n}$ choices for $W$ (because we
can choose the sign of $\varepsilon_{r}$), so $E_{0}/F$ permutes them
transitively and, by conjugating with $e_{2r}e_{2r+1}$, we can change
the sign of $\varepsilon_{r}$ without changing anything else. Really?
Let us prove it:
\begin{calculation}
  (e_{2r}e_{2r+1})(e_{2r-1}e_{2r})(e_{2r}e_{2r+1})^{-1}
\step{since $(e_{2r}e_{2r+1})^{-1} = e_{2r+1}e_{2r}$}
  (e_{2r}e_{2r+1})(e_{2r-1}e_{2r})e_{2r+1}e_{2r}
\step{associativity}
  e_{2r}e_{2r+1}e_{2r-1}(e_{2r}e_{2r+1})e_{2r}
\step{anticommutativity}
  e_{2r}e_{2r+1}e_{2r-1}((-1)e_{2r+1}e_{2r})e_{2r}
\step{associativity, moving $-1$ to the front}
  (-1)e_{2r}e_{2r+1}e_{2r-1}e_{2r+1}(e_{2r}e_{2r})
\step{since $e_{r}^{2}=-1$}
  (-1)e_{2r}e_{2r+1}e_{2r-1}e_{2r+1}(-1)
\step{move $-1$ to the front, associativity}
  (-1)^{2}e_{2r}(e_{2r+1}e_{2r-1})e_{2r+1}
\step{anticommutativity}
  (-1)^{2}e_{2r}((-1)e_{2r-1}e_{2r+1})e_{2r+1}
\step{move $-1$ to the front, associativity}
  (-1)^{3}e_{2r}e_{2r-1}(e_{2r+1}e_{2r+1})
\step{since $e_{2r+1}^{2}=-1$}
  (-1)^{3}e_{2r}e_{2r-1}(-1)
\step{move $-1$ to the front, anticommutativity}
  (-1)^{4}(-e_{2r-1}e_{2r})
\step{since $(-1)^{4}=1$}
  -e_{2r-1}e_{2r}.
\end{calculation}
This is precisely what Adams claimed.

Each of these possible choices for $\varepsilon_{r}=\pm1$ appears in
$\CC[E_{0}]\otimes_{\CC[F]}W$. Therefore, we get a 
representation $\Delta$ of $E_{0}$ with characters $2^{n}$ at $1$,
$-2^{n}$ at $-1$, and $0$ everywhere else. By orthogonality relations,
$\Delta$ is an irreducible representation of $E_{0}$.

\medbreak
\textbf{Case 2: $m=2n$}. There are $2^{n}$ choices for $W$  and under
$E_{0}/F$ falls into two orbits: those with $\prod\varepsilon_{r}=+1$
and those with $\prod\varepsilon_{r}=-1$. We define $\varepsilon:=\prod\varepsilon_{r}$.

We can change the sign of an even number of $\varepsilon_{r}$: by
conjugating by $e_{2r}e_{2s}$ for $r<s$, we change the sign for both
$\varepsilon_{r}$ and $\varepsilon_{s}$. Really? Well, let us prove it:
\begin{calculation}
  e_{2r}e_{2s}(e_{2r-1}e_{2r})(e_{2r}e_{2s})^{-1}
\step{since $(e_{2r}e_{2s})^{-1}=e_{2s}e_{2r}$}
  e_{2r}e_{2s}(e_{2r-1}e_{2r})e_{2s}e_{2r}
\step{associativity}
  e_{2r}e_{2s}e_{2r-1}(e_{2r}e_{2s})e_{2r}
\step{anticommutativity}
  e_{2r}e_{2s}e_{2r-1}((-1)e_{2s}e_{2r})e_{2r}
\step{associativity}
  e_{2r}e_{2s}e_{2r-1}(-1)e_{2s}(e_{2r}e_{2r})
\step{since $e_{2r}^{2}=-1$}
  e_{2r}e_{2s}e_{2r-1}(-1)e_{2s}(-1)
\step{move $-1$ to the front, associativity}
  (-1)^{2}e_{2r}e_{2s}(e_{2r-1}e_{2s})
\step{anticommutativity}
  (-1)^{2}e_{2r}e_{2s}((-1)e_{2s}e_{2r-1})
\step{move $-1$ to the front, associativity}
  (-1)^{3}e_{2r}(e_{2s}e_{2s})e_{2r-1}
\step{since $e_{2s}^{2}=-1$}
  (-1)^{3}e_{2r}(-1)e_{2r-1}
\step{move $-1$ to the front}
  (-1)^{4}e_{2r}e_{2r-1}
\step{anticommutativity}
  (-1)^{5}e_{2r-1}e_{2r} = -e_{2r-1}e_{2r}.
\end{calculation}
Similar calculation works for $e_{2s-1}e_{2s}$.

The character for these representations is $2^{n-1}$ at $1$, $-2^{n-1}$
at $-1$, $\I^{n}\varepsilon$ at $\prod^{2n}e_{i}$, and $-\I^{n}\varepsilon$
at $-\prod^{2n}e_{i}$, and $0$ everywhere else. These are irreducible
representations by orthogonality relations. These are also inequivalent
representations denoted $\Delta^{\pm}$, which concludes the proof.

(Also, observe from Lemma~\ref{lemma:spin:center-of-e0} that
$\prod^{2n}e_{i}$ lives in the center of $E_{0}$ and must commute with
everything. This implies it is represented by a diagonal matrix, and I
think $\Delta^{\pm}$ are the eigenspaces for it.)
\end{proof}

\begin{remark}[Note to self]
If I were to revise these notes, it might be wise to turn the
calculations ``Conjugate $\varepsilon_{r}$ by these quadratic terms''
into exercises.
\end{remark}

\begin{remark}
If we were to think of the induced representation as a hom-set (\S\ref{chunk:spin:induced-rep-as-hom-set}),
then $W\iso\{f|_{\{1\}}\mid f\in\hom_{F}(G,\CC)\}$ I think (since
$W\iso\CC^{1}$ is a one-dimensional space).
\end{remark}

\begin{proposition}[{Adams~\cite[Prop.4.4]{adams1996:ex}}]\label{prop:spin:spinor-rep-inclusion}
Let $\Delta^{\pm}_{2n}$ be the irreps for $\Spin(2n)$, $\Delta_{2n+1}$
the irrep for $\Spin(2n+1)$, and $\Delta^{\pm}_{2n+2}$ the irreps for
$\Spin(2n+2)$. Then under the obvious inclusions
$\Spin(2n)\into\Spin(2n+1)\into\Spin(2n+2)$, we have the representations
restrict as
\[ \Delta^{+}_{2n}\oplus\Delta^{-}_{2n} \iso \Delta_{2n+1}|_{\Spin(2n)} \]
and
\[ \Delta_{2n+1} \iso \Delta^{\pm}_{2n+2}|_{\Spin(2n+1)}. \]
\end{proposition}

The more general situation --- relating (1) irreps of $\Spin(r)$, $\Spin(s)$
in the product group $\Spin(r)\times\Spin(s)$ and (2) irreps of $\Spin(r+s)$
--- are handled in the next proposition.

\begin{proposition}[{Adams~\cite[Prop.4.5]{adams1996:ex}}]
\begin{enumerate}
\item Under the ``obvious'' inclusion group morphism
  $\Spin(2r)\times\Spin(2s)\into\Spin(2r+2s)$, we have the retraction
  [left inverse morphism] induce mappings of the representations when
  restricted to the subgroups:
  \[ \Delta_{2r}^{+}\otimes\Delta_{2s}^{+} + \Delta^{-}_{2r}\otimes\Delta^{-}_{2s}
\gets \Delta_{2r+2s}^{+}|_{\Spin(2r)\times\Spin(2s)} \]
  \[ \Delta_{2r}^{+}\otimes\Delta_{2s}^{-} + \Delta^{-}_{2r}\otimes\Delta^{+}_{2s}
\gets \Delta_{2r+2s}^{-}|_{\Spin(2r)\times\Spin(2s)} \]
\item Under $\Spin(2r)\times\Spin(2s + 1)\into\Spin(2r+2s + 1)$, we have
  \[ \Delta_{2r}^{+}\otimes\Delta_{2s+1} + \Delta^{-}_{2r}\otimes\Delta_{2s+1}
\gets \Delta_{2r+2s+1}|_{\Spin(2r)\times\Spin(2s+1)} \]
\item Under $\Spin(2r+1)\times\Spin(2s+1)\into\Spin(2r+2s+2)$, we have
  \[ \Delta_{2r+1}\otimes\Delta_{2s+1}
\gets \Delta_{2r+2s+2}^{+}|_{\Spin(2r+1)\times\Spin(2s+1)} \]
  \[ \Delta_{2r+1}\otimes\Delta_{2s+1}
\gets \Delta_{2r+2s+2}^{-}|_{\Spin(2r+1)\times\Spin(2s+1)} \]
\end{enumerate}
\end{proposition}

I'll just sketch the proof scheme which Adams uses.

\begin{proof}[Proof scheme]
The weights agree with the modules on the left and on the right, which
implies the characters agree on the compact group $\Spin_{\RR}(p)\times\Spin_{\RR}(q)$, and establishes the modules are
isomorphic. Let us work through the first claim's proof explicitly, so
$p=2r$ and $q=2s$.

For example, if we take a weight
$\frac{1}{2}(\varepsilon_{1}x_{1}+\cdots+\varepsilon_{r+s}x_{r+s})$ with
$\varepsilon_{1}\varepsilon_{2}\cdots\varepsilon_{r+s}=1$, then we get
on the right
$\frac{1}{2}(\varepsilon_{1}x_{1}+\cdots+\varepsilon_{r}x_{r}) + \frac{1}{2}(\varepsilon_{r+1}x_{r+1}+\cdots+\varepsilon_{r+s}x_{r+s})$
with either $\varepsilon_{1}\cdots\varepsilon_{r}=1$ and 
$\varepsilon_{r+1}\cdots\varepsilon_{r+s}=1$
or $\varepsilon_{1}\cdots\varepsilon_{r}=-1$ and 
$\varepsilon_{r+1}\cdots\varepsilon_{r+s}=-1$. This shows the characters
agree on the product group $\Spin_{\RR}(2r)\times\Spin_{\RR}(2s)$.

Then the characters agree on the finite groups $E_{0}(2r)\times E_{0}(2s)$.
Then the modules are isomorphic as representations of these finite
groups, hence isomorphic as modules over $\FF[E_{0}(2r)\times E_{0}(2s)]$.

But they are isomorphic as modules over
$\bigl(\FF[E_{0}(2r)]/(\nu+1)\bigr)\otimes\bigl(\FF[E_{0}(2s)]/(\nu+1)\bigr)\iso\Cl^{0}(2r)\otimes_{\FF}\Cl^{0}(2s)$.
So the modules are isomorphic as representations of
$\Spin_{\FF}(2r)\times\Spin_{\FF}(2s)$, where
$\Spin_{\FF}(p)=\Spin(\FF^{p})$ for $\FF=\RR$ or $\FF=\CC$.
\end{proof}

\subsection{Fermionic Operator Algebra Versions of Half-Spin Representations}


\M
Consider $\Spin(m)$ for $m=2n$. We will construct a complex Hilbert space
$\mathcal{H}$ equipped with constructor operators $a^{\dagger}_{i}$ and
destruction operators $a_{i}$, for $i=1,\dots,n$. They satisfy the
canonical anticommutation relations
\begin{subequations}
\begin{align}
\{a_{i},a_{j}\} = \{a^{\dagger}_{i},a^{\dagger}_{j}\} = 0\\
\{a_{i},a^{\dagger}_{j}\} = \delta_{i,j}\mathbf{1}
\end{align}
\end{subequations}
where $i$, $j=1,$ \dots, $n$, and $\mathbf{1}$ is the
identity operator acting on $\mathcal{H}$. 
We also have a vacuum vector $|0\rangle\in\mathcal{H}$ which is
annihilated by all creation operators
\begin{equation}
a_{i}|0\rangle = 0.
\end{equation}
Furthermore, the vacuum vector is not the zero vector
\begin{equation}
|0\rangle\neq 0.
\end{equation}

\begin{ddanger}
This only works for $\Spin(2n)$ groups, with $m=2n$ even. I cannot
stress this enough. For $\Spin(2n+1)$, this will not work
exactly. Instead, we will relate this fermionic operator algebra to the
usual generators of the Clifford algebra, and this will relate the
half-spin representations to complementary subspaces of $\mathcal{H}$.
\end{ddanger}

\begin{remark}
We observe that we can form a representation of these guys using the
Grassmann algebra $\CC^{0|n}$. Then the vacuum vector is represented by
$1$, the creation operators $a^{\dagger}_{i}$ are represented by odd
variables $\theta_{i}$, the destruction operators $a_{i}$ are the
derivatives $\partial/\partial\theta_{i}$. Not only is this quite nifty,
it's \emph{super} nifty.
\end{remark}

\N{Clifford Generators}
We can recover the generators for the Clifford algebra by
\begin{equation}
a_{j} = e_{2j-1} + \I e_{2j},\quad\mbox{and}\quad
a^{\dagger}_{j} = \frac{1}{\I}(e_{2j-1} - \I e_{2j}).
\end{equation}
Then
\begin{equation}
e_{2j-1} = \frac{a_{j} - \I a^{\dagger}_{j}}{2},\quad\mbox{and}\quad
e_{2j} = \frac{a_{j} + \I a^{\dagger}_{j}}{2\I}.
\end{equation}
Observe then that
\begin{equation}
4\I e_{2j-1}e_{2j} = a_{j}^{2} - \I a_{j}^{\dagger}a_{j} +
\I a_{j}a_{j}^{\dagger} + (a_{j}^{\dagger})^{2}.
\end{equation}
The square of creation operators vanish, as do the square of
annihilation operators. We see then that the right-hand side is equal to
$\pm2\I\mathbf{1}$, which relates back to Adams's construction of the
irreps for $\Spin(2n)$, since these are representations of the Abelian
subgroup $F<E_{0}$ generated by $\pm e_{2j-1}e_{2j}$.

\N{Half-Spin Representations}
We then have for $m=2n$ two subrepresentations, namely all powers of
quadratic combinations of construction operators acting on the vacuum
\begin{subequations}
\begin{align}
S^{+} = \operatorname{Span}\{|0\rangle, a^{\dagger}_{i}a^{\dagger}_{j}|0\rangle,\dots,a^{\dagger}_{1}a^{\dagger}_{2}(\cdots)a^{\dagger}_{n}|0\rangle\}
\intertext{and}
S^{-} = \operatorname{Span}\{a^{\dagger}_{i}|0\rangle, a^{\dagger}_{i}a^{\dagger}_{j}a^{\dagger}_{k}|0\rangle,\dots\}
\end{align}
\end{subequations}
where $1\leq i<j<k<\dots\leq n$. These correspond to $\Delta^{+}$, $\Delta^{-}$
(respectively).

\M
The fermionic operator algebra gives us irreducible representations for
$\spin(2m)$. Exponentiation then gives us irreducible representations
for the group $\Spin(2m)$. But we are required to exponentiate things.

\subsection{Tensor Products of Spinor Representations}

\N{Exterior powers of fundamental representations of $\SO(2n)$}
For $\SO(m)$ we can describe its fundamental representation
$V\iso\FF^{m}$ as well as its exterior powers $\Exterior^{k}V$. There is
something special that happens for the case when $m=2n$ and $k=n$,
namely we get two irreducible subrepresentations:
$\Exterior^{n}V\iso\Exterior^{n+}V\oplus\Exterior^{n-}V$.
How does this work?

Well, consider the situation for $V\iso\RR^{4}$ with canonical basis
$e_{1}$, $e_{2}$, $e_{3}$, $e_{4}$. We have the Hodge star:
\begin{subequations}
\begin{align}
\star(e_{1}\wedge e_{2}) &= e_{3}\wedge e_{4}\\
\star(e_{1}\wedge e_{3}) &= -e_{2}\wedge e_{4}\\
\star(e_{1}\wedge e_{4}) &= e_{2}\wedge e_{3}\\
\star(e_{2}\wedge e_{3}) &= e_{1}\wedge e_{4}\\
\star(e_{2}\wedge e_{4}) &= -e_{1}\wedge e_{3}\\
\star(e_{3}\wedge e_{4}) &= e_{3}\wedge e_{4}.
\end{align}
\end{subequations}
This is because $e_{i}\wedge e_{j}\wedge\star(e_{i}\wedge e_{j})=\omega$
is precisely the volume form $\omega = e_{1}\wedge e_{2}\wedge e_{3}\wedge e_{4}$.

Now, the Hodge star gives us two eigenspaces. The $+1$ eigenspace are
spanned by $e_{1}\wedge e_{2} + e_{3}\wedge e_{4}$,
$e_{1}\wedge e_{3} -e_{2}\wedge e_{4}$,
and $e_{1}\wedge e_{4} + e_{2}\wedge e_{3}$.
The reader can verify $\star$ applied to these 2-forms are precisely
themselves. This span describes $\Exterior^{2+}\RR^{4}$.

The $-1$ eigenspace for the Hodge star is spanned by
$e_{1}\wedge e_{2} - e_{3}\wedge e_{4}$,
$e_{1}\wedge e_{3} + e_{2}\wedge e_{4}$, and
$e_{1}\wedge e_{4} - e_{2}\wedge e_{3}$. This span describes
$\Exterior^{2-}\RR^{4}$. The reader can verify these two forms $v$
satisfy $\star v = - v$.

\begin{remark}
The general scheme works for the more general setting of $n$-forms. We
just take $\star (e_{i_{1}}\wedge\dots\wedge e_{i_{n}}) = s e_{j_{1}}\wedge\dots\wedge e_{j_{n}}$
where $s=\pm1$, $i_{1}<\dots<i_{n}$ and $j_{1}<\dots<j_{n}$ and
$\{i_{1},\dots,i_{n},j_{1},\dots,j_{n}\} = \{1,2,\dots,2n\}$.
Then we determine the sign $s$ by computing
\begin{equation}
(e_{i_{1}}\wedge\dots\wedge e_{i_{n}})\wedge\star(e_{i_{1}}\wedge\dots\wedge e_{i_{n}})=\omega.
\end{equation}
Once we have done this for every possible basis $n$-form for
$\Exterior^{n}\FF^{2n}$, we form the eigenspaces:
\begin{subequations}
\begin{align}
\Exterior^{n+}\FF^{2n}=\span\{(e_{i_{1}}\wedge\dots\wedge e_{i_{n}}) + \star (e_{i_{1}}\wedge\dots\wedge e_{i_{n}})\}\\
\intertext{and}
\Exterior^{n-}\FF^{2n}=\span\{(e_{i_{1}}\wedge\dots\wedge e_{i_{n}}) - \star (e_{i_{1}}\wedge\dots\wedge e_{i_{n}})\}.
\end{align}
\end{subequations}
For indefinite metrics, $\RR^{p,q}$, care must be taken for the signs
--- namely, for $q$ odd, the spans describe the opposite eigenspaces.
\end{remark}

\begin{remark}
Thank goodness Qiaochu Yuan clarified this for me.\footnote{He was
answering someone else's question, though, but I still appreciate it! \url{https://mathoverflow.net/a/440634}}
\end{remark}

\N{Notation}
We write: $V+W$ for $V\oplus W$; and
$\Sym^{2}(V)$ is the second symmetric power of $V$. We recall
$V\otimes V\iso\Sym^{2}(V) + \Exterior^{2}V$.

\begin{theorem}[{Adams~\cite[Th.4.6(ii)]{adams1996:ex}}]
Let $V$ be the fundamental representation for $\SO(2n+1)$. As
representations for $\Spin(2n+1)$, we have the following tensor products:
\begin{enumerate}
\item $\Exterior^{2}\Delta\iso\Exterior^{n}V + (\Exterior^{n-3}V + \Exterior^{n-4}V) +\dots+(\Exterior^{n-3-4k}V + \Exterior^{n-4-4k}V)+\dots$
\item $\Sym^{2}(\Delta)\iso(\Exterior^{n-1}V + \Exterior^{n-2}V) + (\Exterior^{n-5}V + \Exterior^{n-6}V) + \dots + (\Exterior^{n-1-4k}V + \Exterior^{n-2-4k}V) +\dots$.
\end{enumerate}
\end{theorem}

\begin{proof}[Proof sketch]
  It follows from Schur's lemma that
  \begin{equation}
\Delta\otimes\Delta\iso\bigoplus^{n}_{j=0}\Exterior^{2j}V^{*}\iso\bigoplus^{n}_{j=0}\Exterior^{2j}V.
  \end{equation}
  We split this sum into two parts --- the summands whose exterior
  powers are not greater than $n$, and the summands whose exterior
  powers are strictly greater than $n$
\begin{equation}
\Delta\otimes\Delta
\iso\left(\bigoplus^{n/2}_{j=0}\Exterior^{2j}V\right)
\oplus\left(\bigoplus^{n}_{j=n/2}\Exterior^{2j}V\right).
\end{equation}
The bounds are chosen such that the first term includes $\Exterior^{n}V$.
For $2j>n$, we take the Hodge star to write this as:
\begin{equation}
\Delta\otimes\Delta
\iso\left(\bigoplus^{n/2}_{j=0}\Exterior^{2j}V\right)
\oplus\left(\bigoplus^{n}_{j=n/2}\Exterior^{2n+1-2j}V\right).
\end{equation}
We can reindex the odd exterior powers, which gives us the isomorphism:
  \begin{equation}\label{eq:spin:delta-otimes-delta-equals-bigoplus}
\Delta\otimes\Delta\iso\bigoplus^{n}_{j=0}\Exterior^{j}V.
  \end{equation}
  We then use the fact that
  \begin{equation}
\Delta\otimes\Delta\iso\Sym^{2}(\Delta)\oplus\Exterior^{2}\Delta.
  \end{equation}
  The result follows.
\end{proof}

\begin{remark}
The use of the Hodge star is responsible for the shifted indices.
\end{remark}

\begin{remark}
Adams uses the canonical injection of $V^{\otimes k}$ into $\Cl(V)$
by
\begin{equation}
\vec{v}_{1}\otimes\dots\otimes\vec{v}_{k}\mapsto\sum_{\sigma\in S_{k}}\frac{\sgn(\sigma)}{k!}v_{\sigma(1)}\otimes\dots\otimes v_{\sigma(k)}.
\end{equation}
This leads to vector space isomorphisms $\oplus\Exterior^{2k}V\to\Cl^{0}(V)$.
(We only need vector space isomorphisms, because these are isomorphisms
of the representations spaces. We also only need the even part of
$\Cl(V)$, which corresponds to the even part of
$\Exterior V$. This works for any field $\FF$ whose character is not 2.)
Adams uses this to obtain $\Delta\otimes\Delta\iso\Exterior^{0}V + \Exterior^{2}V + \Exterior^{4}V + \dots + \Exterior^{2n}V$.
From there, he uses the Hodge star to obtain Eq~\eqref{eq:spin:delta-otimes-delta-equals-bigoplus}.
The rest of the proof remains the same.
\end{remark}

\begin{theorem}[{Adams~\cite[Th.4.6(i)]{adams1996:ex}}]
Let $V$ be the fundamental representation for $\SO(2n)$. As
representations for $\Spin(2n)$, we have the following tensor products:
\begin{enumerate}
\item $\Sym^{2}(\Delta^{+}) \iso \Exterior^{n+}V + \Exterior^{n-4}V + \Exterior^{n-8}V+\dots$
\item $\Exterior^{2}\Delta^{+} \iso \Exterior^{n-2}V + \Exterior^{n-6}V + \dots$
\item $\Sym^{2}(\Delta^{-}) \iso \Exterior^{n-}V + \Exterior^{n-4}V + \Exterior^{n-8}V+\dots$
\item $\Exterior^{2}\Delta^{-} \iso \Exterior^{n-2}V + \Exterior^{n-6}V + \dots$
\item $\Delta^{+}\otimes\Delta^{-}\iso\Exterior^{n-1}V + \Exterior^{n-3}V + \Exterior^{n-5}V + \dots$.
\end{enumerate}
\end{theorem}

The proof is similar to the previous one, taking advantage of
Proposition~\ref{prop:spin:spinor-rep-inclusion}. Adams proves
the $\Delta^{+}\otimes\Delta^{-}$ isomorphism, the other cases being
``obvious''. 

\section{Miscellaneous}

\M
The basic idea is that, when $m=2n+1$ is odd, then
$\omega\notin\Cl^{0}(m)$. Therefore,
\begin{equation}
\Cl(m) = \Cl^{0}(m) \oplus \omega\,\Cl^{0}(m)\iso\Cl^{0}(m)\otimes_{\RR}\RR[\omega].
\end{equation}
We can induce a Pinor representation $P$ of $\Cl^{0}(m)$ from a Spinor
representation $S$ of $\Cl^{0}(m)$ by tensoring it
with $\RR[\omega]$:
\begin{equation}
P = \Cl(m)\otimes_{\Cl^{0}(m)} S.
\end{equation}
Observe
\begin{equation}
\RR[\omega]\iso\CC\quad\mbox{when}\quad m=1,5\pmod8
\end{equation}
since $\omega^{2}=-1$ in that case; and
\begin{equation}
\RR[\omega]\iso\RR\oplus\RR\quad\mbox{when}\quad m=3,7\pmod8
\end{equation}
since $\omega^{2}=1$ in those cases.

\N{``Definition''}
In even dimensions, there are two inequivalent irreducible
representations for the $\Spin(2n)$ group. Physicists call them
\define{Weyl Spinors}. The volume element $\omega$ plays the role of
picking out the irreducible representations; more precisely
$(1\pm\omega)$ projects onto the two irreducible representations.

\M
There are two inequivalent complex spinor representations when $s-t=2,6\pmod8$
namely $S(s,t)$ and $\overline{S(s,t)}$ of complex dimension
$2^{(s+t-2)/2}$. They are determined by the value of $\Gamma_{d+1}$: it is
$i$ on $S(s,t)$ and $-i$ on $\overline{S(s,t)}$.

\N{Majorana--Weyl Spinors}
When $s-t=0\pmod8$, there are two inequivalent real spinor
representations $S(s,t)_{\pm}$ of real dimension $2^{(s+t-2)/2}$ and
distinguished by the value of $\Gamma_{d+1}$: being $\pm1$ on $S(s,t)_{\pm}$.
These are known as the Majorana--Weyl Spinors.

\N{Symplectic Majorana--Weyl Spinors}
When $s-t=4\pmod8$, there are two inequivalent quaternionic spinor
representations $S(s,t)_{\pm}$ of quaternionic dimension $2^{(s+t-4)/2}$
and distinguished by the value of $\Gamma_{d+1}$ being $\pm1$ on
$S(s,t)_{\pm}$.
These are the \emph{symplectic Majorana--Weyl Spinors}.

\N{References}
See also \S3 of Li's ``Geometry of Supergravity''\footnote{\url{https://sili-math.github.io/Supersymmetry.pdf}},
Todorov's ``Clifford Algebras and Spinors''\footnote{\url{https://inis.iaea.org/collection/NCLCollectionStore/_Public/43/048/43048770.pdf?r=1&r=1}}.


\begin{example}
\begin{enumerate}
\item $\Spin(2)\iso\U(1)$ is a circle
\item $\Spin(3)\iso\SU(2)\iso\Sp(1)$ using $S^{3}\subset\HH$
\item $\Spin(4)\iso\SU(2)\times\SU(2)$ using two copies of $S^{3}\subset\HH$
\item $\Spin(5)\iso\Sp(2)$ corresponds to unitary group of quaternionic
  $2\times2$ matrices
\end{enumerate}
\end{example}

\M
Adams then uses these representations to construct exceptional Lie
groups. For example, he constructs [the compact real] $\mathtt{G}_{2}$
as a subgroup of $\Spin(7)$ in Adams's Theorem~5.5.

Its split form may be constructed analogously, but using 7-dimensional
Minkowski space instead.
Gogberashvili and Gurchumelia~\cite{Gogberashvili:2019ojg} construct
this quite explicitly. Ekins and Cornwell~\cite{Ekins:1975yu} do the
calculations for the associated Lie algebra.

\M
Also, Adams makes heavy usage of the $\Spin(16)$ subgroup of
$\mathtt{E}_{8}$ when constructing other exceptional Lie groups.
We have a sequence of mappings $G\to\Spin(16)\to\mathtt{E}_{8}$,
then take the centralizer of the image of $G$ under these maps, and the
identity component of that centralizer turns out to be a compact form of
an exceptional Lie group (depending on the choice of $G$).
This provides compact forms for $\mathtt{E}_{6}$, $\mathtt{E}_{7}$,
$\mathtt{F}_{4}$.

The only direct construction for $\mathtt{F}_{4}$
that I can find is Figueroa-O'Farrill~\cite{Figueroa-OFarrill:2007jcv}
using $S^{8}$ and Killing spinors.
