\section{Poincar\'e Algebra}

\N{Big Idea}
We saw in Section~\ref{section:relativity:group-theoretic-description}
how to ``implement'' special relativity using group
theory. Specifically, we saw how the Poincar\'e group encodes special
relativity. 

We study Lie groups by studying their Lie algebras (``infinitesimal
symmetries near the identity''), and we want to find finite-dimensional
unitary representations for the Poincar\'e group. It suffices for our
purposes to study the Lie algebra for the Lorentz group (which can be
extended to the Lie algebra for the Poincar\'e group by adjoining
generators of translations --- i.e., the momentum and Hamiltonian
operators). We want to find irreducible representations, since these
correspond to ``elementary'' quantum systems.

Wigner forwarded the interpretation that irreducible representations for
the ``Poincar\'e group'' should be viewed as elementary
particles.\index{Elementary Particle!As Irreducible Representation}

\subsection{Lorentz Algebra}

\begin{definition}
The subgroup of \define*{Proper Orthochronous Lorentz Transformations}\index{Lorentz transformation!Proper orthochronous}
denoted either as $\ISO(3,1)$ or $\SO^{+}(3,1)$, consists of Lorentz
transformations $\Lambda\in\O(3,1)$ such that
\begin{enumerate}
\item it is proper (``preserves orientation''): $\det(\Lambda)=+1$, and
\item it is orthochronous: ${\Lambda^{0}}_{0}\geq+1$.
\end{enumerate}
\end{definition}
\begin{exercise}
Prove the identity transformation is a proper orthochronous Lorentz
transformation. 
\end{exercise}

\N{Non-examples}
There are two critical example:
\begin{enumerate}
\item The transformation of time reversal $T=\diag(-1,1,1,1)$ is not an
  orthochronous Lorentz transformation (but it is a Lorentz transformation).
\item Parity reversal $P=\diag(1,-1,-1,-1)$ is a Lorentz transformation
  but it is not proper.
\end{enumerate}
Composing these two operators does not give us a proper orthochronous
Lorentz transformation. But in four dimensions, the Lorentz group
$\O(3,1)$ has four connected components, and applying $P$, $T$, $PT$ to
the identity component gives us the other three. Therefore, we can work
with an arbitrary proper orthochronous Lorentz transformation
$\Lambda\in\ISO(3,1)$ then $P\Lambda$, $T\Lambda$, $PT\Lambda$ gives us
arbitrary Lorentz transformations \emph{in four dimensions}.

\N{Obtaining the Lorentz Lie Algebra}
It's a common trick to parametrize a proper orthochronous Lorentz
transformation $\Lambda\in\ISO(3,1)$ using 
``angles'' [scalars], just as we would with rotations, as
\begin{equation}
{\Lambda^{\mu}}_{\nu} = [\exp\left(\frac{-\I}{2}\omega_{\kappa\lambda}M^{\kappa\lambda}\right)]{{}^{\mu}}_{\nu}
\end{equation}
where $\omega_{\kappa\lambda}=-\omega_{\lambda\kappa}$ are ``rotation
angles'' (real constants parametrizing the symmetry) and
$M^{\kappa\lambda}$ is an indexed family of matrices (i.e., fix a value
of $\kappa$ and $\lambda$, and you get a $4\times4$ matrix). These $M^{\kappa\lambda}$ are
generators of the Lie algebra for the Lorentz group. Explicitly
\begin{equation}
(M^{\kappa\lambda})_{\mu\nu} = \I(\delta^{\kappa}_{\mu}\delta^{\lambda}_{\nu}-\delta^{\kappa}_{\nu}\delta^{\lambda}_{\mu})
\end{equation}
Now the trick is that we can write the generators of the Lorentz Lie
algebra using
\begin{subequations}\label{eq:rqm:poincare-algebra:rotations-and-boosts-generators}
\begin{align}
L^{i} &= \frac{1}{2}\epsilon^{ijk}M_{jk}\\
\intertext{for spatial rotations, and}
K^{i} &= M^{0i}%\\
%% \intertext{for Lorentz boosts. We define}
%% \vec{J}_{\pm} &= \frac{1}{2}(\vec{L}\pm\I\vec{K}).
\end{align}
\end{subequations}

\N{Commutation Relations}\label{chunk:rqm:poincare-algebra:lorentz-algebra:commutation-relations}
We can obtain the commutation relations by considering an arbitrary
Lorentz transformation $\Lambda\in\ISO(3,1)$ and an infinitesimal
Lorentz transformation $1+\delta\omega$, then evaluating to first order
in $\delta\omega$ gives us:
\begin{subequations}
\begin{align}
  U(\Lambda,0)^{-1}U(1 + \delta\omega,0)U(\Lambda,0)
  &= 1+U(\Lambda,0)^{-1}M^{\mu\nu}U(\Lambda,0)\delta\omega_{\mu\nu}\\
  &= 1+{\Lambda^{\mu}}_{\rho}{\Lambda^{\nu}}_{\sigma}M^{\rho\sigma}\delta\omega_{\mu\nu}.
\end{align}
\end{subequations}
Since $\delta\omega_{\mu\nu}$ is arbitrary (but antisymmetric), we have
\begin{equation}
U(\Lambda,0)^{-1}M^{\mu\nu}U(\Lambda,0)
= {\Lambda^{\mu}}_{\rho}{\Lambda^{\nu}}_{\sigma}M^{\rho\sigma}.
\end{equation}
Now if we let $\Lambda$ be an infinitesimal Lorentz transformation, we
arrive at
\begin{equation}
[M^{\mu\nu}, M^{\rho\sigma}]
= (\eta^{\mu\sigma}M^{\nu\rho} - \eta^{\nu\sigma}M^{\mu\rho})
- (\eta^{\mu\rho}M^{\nu\sigma} - \eta^{\nu\rho}M^{\mu\sigma}).
\end{equation}
This gives us the commutation relations in terms of rotations $L^{i}$
and Lorentz boosts $K^{i}$ as
\begin{subequations}
\begin{align}
[L^{i}, L^{j}] &= \I{\epsilon^{ij}}_{k}J^{k}\\
[L^{i}, K^{j}] &= \I{\epsilon^{ij}}_{k}K^{k}\\
[K^{i}, K^{j}] &= -\I{\epsilon^{ij}}_{k}J^{k}.
\end{align}
\end{subequations}

\begin{exercise}
Double check the calculations are correct. There may be a sign error.
\end{exercise}

\N{Complexified Lorentz Lie Algebra}
We are really interested in the irreducible complex linear
representations of the complexification $\iso(3,1)_{\CC}$ of the Lie
algebra $\iso(3,1)$ of the Lorentz group. This is because the Lorentz
group is non-compact (which makes it impossible to obtain a
finite-dimensional Unitary representation), so we change the goal to
find holomorphic representations of the complexified Lie Algebra for the
Lorentz group. This is precisely Weyl's Unitarian trick.

We define
\begin{equation}
\vec{J}_{\pm} = \frac{1}{2}(\vec{L}\pm\I\vec{K}).
\end{equation}
The reader may verify the commutation relations become
\begin{equation}
[J^{i}_{\pm}, J^{j}_{\pm}] = \I\epsilon^{ijk}J^{k}_{\pm}.
\end{equation}
But now look, this is precisely two copies of the complexified Lie
algebra $\su(2)_{\CC}$ (each copy is isomorphic to $\sl(2,\CC)$)
parametrized by the ``$\pm$'' subscripts.

Thus by the Unitarian trick,\index{Unitarian trick} we can study representations of $\iso(3,1)$
by means of representations for $\su(2)_{\CC}$. Specifically
highest-weight representations of $\su(2)_{\CC}$.

\M\label{chunk:rqm:poincare-algebra:elementary-particles-irreps}
The punchline:
\textit{Each irreducible representation of $\iso(3,1)$
is characterized by a pair of half-integers $(j_{+}, j_{-})$.} We can
interpret these irreducible representations as particles, summarized by
the handy-dandy table:

\begin{center}
\begin{tabular}{c|c|c}
  $(j_{+}, j_{-})$ & Name of Field & Dimension of Rep \\\hline
  $(0, 0)$ &	Scalar  &	1\\
$(1/2, 0)$ & 	Left-handed Weyl Spinor &	2\\
$(0, 1/2)$ &	Right-handed Weyl Spinor &	2\\
$(1, 0)$ &	(Imaginary) Self-dual 2-form &	3\\
$(0, 1)$ &	(Imaginary) Anti-self-dual 2-form &	3\\
$(1/2, 1/2)$ &	Vector (gauge field) &	4\\
$(1/2, 1)$ & 	Left-Handed Rarita-Schwinger field &	6\\
$(1, 1/2)$ &	Right-Handed Rarita-Schwinger field &	6\\
$(1, 1)$ &	Graviton (spin-2 field) &	9
\end{tabular}
\end{center}

Quantum Field Theory studies merely three such representations:
\begin{itemize}
\item $(0,0)$ the scalar particle,
\item $(1/2,0)\oplus(0,1/2)$ Dirac fermionic particle, and
\item $(1/2, 1/2)$ vector (gauge Boson) particles.
\end{itemize}

\begin{ddanger}\index{Spin!Higher than 2}%
There is also a ``no-go theorem'' which states there are no interacting
fields of spin $J>2$. That is to say, the only forces we can expect in
the universe are mediated by Scalars, Vector [gauge] bosons, and Spin-2
bosons. There are additional theorems\index{Spin!Uniqueness of (---)-2} stating all Spin-2 forces couple
to energy in exactly the same way, up to some scaling of the coupling
constant --- that is to say, Spin-2 forces all are gravity up to choice
of units.
\end{ddanger}

\N{Infinite-Dimensional Unitary Representations}
Finite-dimensional representations have Lorentz transformations act on
finite-dimensional constant vectors. If we want to have Lorentz
transformations act on a multi-component field $\Phi_{a}(t,\vec{x})$,
then we need a Unitary infinite-dimensional representation for the
Lorentz group (or Poincar\'e group), something like:
\begin{equation}
\Phi_{a}\xrightarrow{\Lambda}\sum_{b}M_{ab}(\Lambda)\Phi_{b}
\end{equation}
where $M(\Lambda)$ treats the components of the field as components of a
constant vector. This is close but incomplete, we can also transform the
spacetime position:
\begin{equation}
\Phi_{a}(x)\xrightarrow{\Lambda}\sum_{b}M_{ab}(\Lambda)\Phi_{b}(\Lambda^{-1}x).
\end{equation}
As a ``smoke check'', we see that a scalar field is just a function on
spacetime $\psi(x)$, and a representation of a group $G$ on function space
gives
\begin{equation}
\forall g\in G, \rho(g)\psi(x)=\psi(g^{-1}\cdot x)
\end{equation}
where there's an ``obvious'' representation of $G$ on the domain of $\psi$.
And behold: the Poincar\'e group acts in an obvious way on space-time!

% https://sharif.edu/~sadooghi/QFT-I-96-97-2/LorentzPoincareMaciejko.pdf

\subsection{Poincar\'e Algebra}

\M We can adjoin the space-time translation generators to the algebra of
Lorentz transformations. We see $H$ generates time translations and
$P^{j}$ generates spatial translations.

\M
We know
\begin{subequations}\label{eq:rqm:poincare-algebra:conjugation-of-elements}
\begin{align}
U(\Lambda,a)M^{\alpha\beta}U^{-1}(\Lambda,a) 
&= {\Lambda_{\mu}}^{\alpha}{\Lambda_{\nu}}^{\beta}(M^{\mu\nu} - a^{\mu}P^{\nu} + a^{\nu}P^{\mu})\\
U(\Lambda,a)P^{\alpha}U^{-1}(\Lambda,a) &= {\Lambda_{\beta}}^{\alpha}P^{\beta}.
\end{align}
\end{subequations}

\N{Determining Commutation Relations}
We now consider infinitesimal group elements
${\Lambda_{\mu}}^{\nu}={\delta_{\mu}}^{\nu}+{\omega_{\mu}}^{\nu}$ and
$a^{\mu}=\epsilon^{\mu}$ in Eq~\eqref{eq:rqm:poincare-algebra:conjugation-of-elements}.
This gives us
\begin{subequations}
\begin{align}
\I[\frac{1}{2}\omega_{\mu\nu}M^{\mu\nu}-\epsilon_{\mu}P^{\mu},M^{\rho\sigma}]
&={\omega_{\mu}}^{\rho}M^{\mu\sigma} + {\omega_{\nu}}^{\sigma}M^{\rho\nu}
-\epsilon^{\rho}P^{\sigma} + \epsilon^{\sigma}P^{\rho}\\
\I[\frac{1}{2}\omega_{\mu\nu}M^{\mu\nu}-\epsilon_{\mu}P^{\mu},P^{\rho}]
&={\omega_{\mu}}^{\rho}P^{\mu}.
\end{align}
\end{subequations}
Equating coefficients of $\omega_{\mu\nu}$ and $\epsilon_{\mu}$ on both
sides of these equations gives us the commutation rules
\begin{subequations}
\begin{align}
\I[M^{\mu\nu},M^{\rho\sigma}] &= \eta^{\nu\rho}M^{\mu\sigma}
- \eta^{\mu\rho}M^{\nu\sigma} - \eta^{\sigma\mu}M^{\rho\nu} + \eta^{\sigma\tau}M^{\rho\mu}\\
\I[P^{\mu},M^{\rho\sigma}] &= \eta^{\mu\rho}P^{\sigma} - \eta^{\mu\sigma}P^{\rho}\\
[P^{\mu},P^{\nu}] &= 0.
\end{align}
\end{subequations}
This is the Lie algebra for the Poincar\'e group.

\M
We can rewrite the commutation relations using the generators for
rotations $L^{i}$ and Lorentz boosts $K^{i}$ from
Eq~\eqref{eq:rqm:poincare-algebra:rotations-and-boosts-generators}.
We recall
(\S\ref{chunk:rqm:poincare-algebra:lorentz-algebra:commutation-relations})
the commutation rules for the Lorentz Lie algebra with these rules:
\begin{subequations}
\begin{align}
[J^{i}, J^{j}] &= \I{\epsilon^{ij}}_{k}J^{k}\\
[J^{i}, K^{j}] &= \I{\epsilon^{ij}}_{k}K^{k}\\
[K^{i}, K^{j}] &= -\I{\epsilon^{ij}}_{k}J^{k}
\end{align}
\end{subequations}
Now we add the generators $P^{j}$ for spatial translations and $H$ for
temporal translations.
\begin{subequations}
\begin{align}
[J^{i}, P^{j}] &= \I{\epsilon^{ij}}_{k}P^{k}\\
[K^{i}, P^{j}] &= \I\delta^{ij}H\\
[J^{i}, H] &= [P^{i}, H] = [H, H] = 0\\
[K^{i}, H] &= \I P^{i},
\end{align}
\end{subequations}
where the indices $i$, $j$, $k=1,2,3$.

\begin{ddanger}
For our purposes in quantum field theory, we will use the universal
covering group for the Poincar\'e group. The universal covering group
for the Lorentz group $\ISO(3,1)$ is the indefinite Spin group $\Spin(3,1)$.
Since the Poincar\'e group is the semidirect product of translations
$\RR^{3,1}$ with the Lorentz group $\ISO(3,1)$, the universal covering
is the semidirect product with $\Spin(3,1)$. This is referred to as the
\define*{Poincar\'e Spin Group}\index{Poincar\'e!Spin group}\index{Group!Poincar\'e!Spin} or
\emph{Poincar\'e Spinor Group} in the literature, though physicists
seldom care about such sordid details. 
\end{ddanger}

\subsection{Single Particle States}

\begin{problem}
Classify all possible 1-particle states invariant under the (identity
component of the) Poincar\'e group.
\end{problem}

\begin{remark}
The hope is to obtain all possible 1-particle Hilbert spaces
$\mathcal{H}$, so that we may construct the Fock space out of it. This
is how quantum field theory works. But this requires classifying the
possible Hilbert spaces $\mathcal{H}$ for 1-particle states.
\end{remark}

\M
Let us write $U(\Lambda,a)$ for the unitary representation of the
transformation $x\to\Lambda x + a$ acting on the Hilbert space of a
single relativistic quantum particle. 
We can classify all states using eigenvectors of the 4-momentum
operator.

We write $\Psi_{p,\sigma}$ for the state with $p$ labeling the momentum eigenvalue,
\begin{equation}
\widehat{P}^{\mu}\Psi_{p,\sigma}=p^{\mu}\Psi_{p,\sigma},
\end{equation}
and $\sigma$ labeling all other degrees of freedom.

\begin{remark}
The $\sigma$ contains finitely many components, as this is the defining
property of a ``single particle''. Several unbound particles could have
$\sigma$ contain continuously many entries.
\end{remark}

\begin{remark}
The distinction between ``elementary particle'' and ``composite particle''
is meaningless here.
\end{remark}

\N{Translations}
Under translations, we find
\begin{equation}
U(1, a)\Psi_{p,\sigma} = \E^{-\I p_{\mu}a^{\mu}}\Psi_{p,\sigma}.
\end{equation}
We need to see how these states transform under homogeneous Lorentz
transformations. 

\N{Boosts}
We see how a homogeneous Lorentz transformation
$U(\Lambda)=U(\Lambda,0)$ acts to produce a state with momentum
eigenvalue $\Lambda p$:
\begin{calculation}
\widehat{P}^{\mu}U(\Lambda)\Psi_{p,\sigma}
\step{multiplying on left by $1 = U(\Lambda)U(\Lambda)^{-1}$}
[U(\Lambda)U(\Lambda)^{-1}]\widehat{P}^{\mu}U(\Lambda)\Psi_{p,\sigma}
\step{associativity}
U(\Lambda)[U(\Lambda)^{-1}\widehat{P}^{\mu}U(\Lambda)]\Psi_{p,\sigma}
\step{Lorentz transformation of momentum operator}
U(\Lambda)[{(\Lambda^{-1})_{\rho}}^{\mu}\widehat{P}^{\rho}]\Psi_{p,\sigma}
\step{since the (linear combination of) eigenvalues are scalars, we pull
  them in front}
[{(\Lambda^{-1})_{\rho}}^{\mu}p^{\rho}]U(\Lambda)\Psi_{p,\sigma}
\step{definition of inverse of Lorentz transformation}
[{\Lambda^{\mu}}_{\rho}p^{\rho}]U(\Lambda)\Psi_{p,\sigma}
\end{calculation}
Therefore $U(\Lambda)\Psi_{p,\sigma}$ lives in the eigenspace of $\widehat{P}^{\mu}$,
and therefore $\widehat{P}^{\mu}U(\Lambda)\Psi_{p,\sigma}$ is a linear
combination of state vectors $\Psi_{p,\sigma'}$ (summing over $\sigma'$):
\begin{equation}\label{eq:rqm:poincare-algebra:one-particle-states:rep-of-poincare-group}
U(\Lambda)\Psi_{p,\sigma} = \sum_{\sigma'}C_{\sigma',\sigma}(\Lambda, p)\Psi_{\Lambda p,\sigma}.
\end{equation}

\N{Species of Particles}
In general, we may use suitable linear combinations of $\Psi_{p,\sigma}$
to choose the $\sigma$ labels in such a way that the matrix
$C_{\sigma',\sigma}(\Lambda,p)$ is block-diagonal. This means that,
within each block, we obtain a subrepresentation of the Poincar\'e group.
It is natural to identify the states of a specific particle type with
the components of a representation of the Poincar\'e group which is
irreducible (in the sense that we cannot decompose it any further).

\begin{problem}
Our current problem can be re-phrased as studying the coefficients
$C_{\sigma',\sigma}(\Lambda,p)$ in irreducible representations of the
Poincar\'e group.
\end{problem}

\N{Trick: ``Physically Distinct'' Momenta}
There is a trick we will use, namely we can use the proper orthochronous Lorentz
boosts on a ``standard'' 4-momenta $k^{\mu}$ to obtain any arbitrary
4-momentum $p^{\mu}$ for a given value of $\eta_{\mu\nu}p^{\mu}p^{\nu}$
and sign of $p^{0}$. This cuts down the space of possible 1-particle
states to these $k^{\mu}$, since they're related to the 1-particle
states with 4-momentum $p^{\mu} = {L^{\mu}}_{\nu}(p)k^{\nu}$ by
\begin{equation}
\Psi_{p,\sigma} = N(p)U(L(p))\Psi_{k,\sigma},
\end{equation}
where $N(p)$ is just some normalization factor.

Now we have a subspace of state vectors $\Psi_{k,\sigma}$ and we want to
study the subgroup of Lorentz transformations preserving this subspace.

\N{Notation}
In this subsection, we will reserve $k^{\mu}$ for the ``standard momenta 4-vectors''.

\begin{definition}
Let $k^{\mu}$ be a standard momentum 4-vector. We define
\define*{Wigner's Little Group}\index{Wigner's little group|see{Group, Wigner's little}}\index{Group!Wigner's little}
to consist of the Poincar\'e transformations
$W\in\O(3,1)$ such that ${W^{\mu}}_{\nu}k^{\nu}=k^{\mu}$, i.e., $W$
stabilizes $k^{\mu}$.
\end{definition}

\begin{theorem}
  For any element $W$ of Wigner's little group, we have
  \begin{equation}
U(W)\Psi_{k,\sigma} = \sum_{\sigma'}D(W)_{\sigma',\sigma}\Psi_{k,\sigma'}
  \end{equation}
  Moreover, the coefficients $D(W)_{\sigma',\sigma}$ give us a
  representation of Wigner's little group.
\end{theorem}

\begin{corollary}
Given an element of the Lorentz group $\Lambda\in\O(3,1)$ and some
4-momentum $p^{\mu}$, we have a corresponding element of Wigner's little
group given by
\begin{equation}
W(\Lambda,p) = L^{-1}(\Lambda p)\cdot\Lambda\cdot L(p).
\end{equation}
\end{corollary}

\begin{corollary}
Given an element of the Lorentz group $\Lambda\in\O(3,1)$ and some
4-momentum $p^{\mu}$, we have $\Lambda$ act on the state vector
$\Psi_{p,\sigma}$ by
\begin{equation}\label{eq:rqm:poincare-algebra:one-particle-states:action-of-lorentz-transformation-on-state}
U(\Lambda)\Psi_{p,\sigma} = \frac{N(P)}{N(\Lambda p)}\sum_{\sigma'}D_{\sigma',\sigma}(W(\Lambda, p))\Psi_{\Lambda p,\sigma'}.
\end{equation}
That is, aside from the normalization constant pulled in front of the
sum, the coefficients $C_{\sigma',\sigma}$ from Eq~\eqref{eq:rqm:poincare-algebra:one-particle-states:rep-of-poincare-group} are determined by the
coefficients $D_{\sigma',\sigma}$ of the representation of Wigner's
little group.
\end{corollary}

\N{Possible One-Particle States}
The possible one-particle states are determined by the standard momentum
$k^{\mu}=(k^{0},\vec{k})$, giving us the following 6 possibilities (where $M$ is some
mass, and $\mu$ is some energy scale like 1 eV):
\begin{center}
\begin{tabular}{ll|l|c}
    & Parameters & Standard Momentum $k^{\mu}$ & Little Group\\\hline
(1*) & $p^{2}=-M^{2}<0$, $p^{0}>0$ & $(M,0,0,0)$ & $\SO(3)$\\
(2) & $p^{2}=-M^{2}<0$, $p^{0}<0$ & $(-M,0,0,0)$ & $\SO(3)$\\
(3*) & $p^{2}=0$, $p^{0}>0$ & $(\mu,0,0,\mu)$ & $\ISO(2)$\\
(4) & $p^{2}=0$, $p^{0}<0$ & $(-\mu,0,0,\mu)$ & $\ISO(2)$\\
(5) & $p^{2}=M^{2}>0$ & $(0,0,0,M)$ & $\SO(2,1)$\\
(6) & $p^{\mu}=0$ & $(0,0,0,0)$ & $\SO(3,1)$
\end{tabular}
\end{center}
The physically meaningful possibilities are options (1) and (3) (and
possibly the vacuum state (6), but that doesn't really describe any particles).
We typically adopt the convention that the $z$-axis is the direction of motion.

Our goal now is to study irreducible representations of Wigner's little
group on these two cases --- the positive-definite mass case, and the
massless case. The massless case will be the relevant one for the
Standard Model.

\begin{ddanger}
We usually interpret these one-particle states as the possible
elementary particles. But we should bear in mind the assumption that
special relativity holds. When we work in curved spacetime, it's no
longer quite so clear what an elementary particle \emph{is}.
\end{ddanger}

\subsection*{Massive Case}

\N{Unitary Representation}
The unitary representation for $\SO(3)$ can be broken up into a direct
sum of irreducible unitary representations $D^{(j)}_{\sigma',\sigma}(R)$
of dimension $2j+1$ for $j=0,\frac{1}{2},1,\frac{3}{2},\dots$ (i.e.,
$j\in\frac{1}{2}\NN_{0}$).

\N{Infinitesimal rotations}
The unitary representations for infinitesimal rotations
$R_{ik}=\delta_{ik}+\Theta_{ik}$ with $\Theta_{ik}=-\Theta_{ki}$
infinitesimal:
\begin{subequations}
\begin{align}
D^{(j)}_{\sigma',\sigma}(1 + \Theta) &= \delta_{\sigma',\sigma} + \frac{\I}{2}\Theta_{ik}(J^{(j)}_{ik})_{\sigma',\sigma}\\
(J^{(j)}_{23}\pm\I J^{(j)}_{31})_{\sigma',\sigma} &= (J^{(j)}_{1}\pm\I J^{(j)}_{2})_{\sigma',\sigma}\nonumber\\
&= \delta_{\sigma',\sigma\pm1}\sqrt{(j\mp\sigma)(j\pm\sigma+1)}\\
(J^{(j)}_{12})_{\sigma',\sigma} &= (J^{(j)}_{3})_{\sigma',\sigma} = \sigma\delta_{\sigma',\sigma},
\end{align}
\end{subequations}
with $\sigma$ running over the values $j$, $j-1$, \dots, $-j$.

\M
For a particle of mass $M>0$ and spin $j$, Eq~\eqref{eq:rqm:poincare-algebra:one-particle-states:action-of-lorentz-transformation-on-state}
becomes
\begin{equation}
U(\Lambda)\Psi_{p,\sigma}
=\sqrt{\frac{(\Lambda p)^{0}}{p^{0}}}\sum_{\sigma'}D^{(j)}_{\sigma',\sigma}(W(\Lambda,p))
\Psi_{\Lambda p,\sigma'}
\end{equation}
with the little-group element $W(\Lambda, p)$ (the Wigner rotation)
given by
\begin{equation*}
W(\Lambda, p) = L^{-1}(\Lambda p)\Lambda L(p).
\end{equation*}
We can calculate this rotation by using the ``standard'' boost $L(p)$
carrying the 4-momentum from $k^{\mu}=(M,\vec{0})$ to $p^{\mu}$. This is
conveniently chosen as:
\begin{subequations}
\begin{align}
{L^{i}}_{k}(p) &= {\delta^{i}}_{k} + (\gamma - 1)\widehat{p}_{i}\widehat{p}_{k}\\
{L^{i}}_{0}(p) &= {L^{0}}_{i}(p) = \widehat{p}_{i}\sqrt{\gamma^{2}-1}\\
{L^{0}}_{0}(p) &= \gamma
\end{align}
\end{subequations}
where
\begin{equation}
\widehat{p}_{i}:=\frac{p_{i}}{\|\vec{p}\|},\quad\mbox{and}\quad
\gamma:=\frac{\sqrt{\|\vec{p}\|^{2}+M^{2}}}{M}.
\end{equation}

\begin{exercise}
Prove that ${L^{\mu}}_{\nu}(p)k^{\nu}=p^{\mu}$.
\end{exercise}

\begin{proposition}
For any three-dimensional rotation $\mathcal{R}$, when the Lorentz
transformation $\Lambda=\mathcal{R}$, the Wigner rotation $W(\mathcal{R},p)=\mathcal{R}$.
\end{proposition}

\N{Consequence}
The states of a moving massive particle have the same transformation
under rotations as in non-relativistic quantum mechanics. We can
therefore use all the tools from non-relativistic quantum mechanics ---
the spherical harmonics, Clebsch-Gordon coefficients, etc. --- to
describe relativistic quantum mechanics for massive particles.

\subsection*{Massless Case}

\begin{theorem}\label{thm:rqm:poincare-algebra:massless-case:general-element-of-little-group}
  The most general element of the little group is of the form
\begin{subequations}
  \begin{equation}
W(\theta,\alpha,\beta) = S(\alpha,\beta)R(\theta),
  \end{equation}
  where
  \begin{equation}
S(\alpha,\beta) = \begin{pmatrix}
1 + \zeta & \alpha & \beta & -\zeta\\
\alpha    &      1 &     0 & -\alpha\\
\beta     &      0 &     1 & -\beta\\
\zeta     & \alpha & \beta & 1 - \zeta
\end{pmatrix},
  \end{equation}
  $\alpha\in\RR$ and $\beta\in\RR$,
 $\zeta = (\alpha^{2} + \beta^{2})/2$,  and
  \begin{equation}
R(\theta) = \begin{pmatrix}
   1 &             0 &            0 & 0\\
     &  \cos(\theta) & \sin(\theta) & 0\\
     & -\sin(\theta) & \cos(\theta) & 0\\
   0 &             0 &            0 & 1
\end{pmatrix}.
  \end{equation}
\end{subequations}
\end{theorem}

\begin{proposition}
We have
\begin{subequations}
\begin{align}
S(\alpha_{1},\beta_{1})S(\alpha_{2},\beta_{2}) &= S(\alpha_{1}+\alpha_{2},\beta_{1}+\beta_{2})\\
R(\theta_{1})R(\theta_{2}) &= R(\theta_{1}+\theta_{2}).
\end{align}
\end{subequations}
In particular, the subgroup generated by the $R(\theta)$ is an Abelian
group, and separately the subgroup generated by the $S(\alpha,\beta)$ is
also Abelian.
\end{proposition}

\begin{theorem}
The Wigner little group is $\ISO(2)$ consisting of translations by a
vector $(\alpha,\beta)$ and rotations by $\theta$.
\end{theorem}

This follows from the previous proposition and the observation that
\begin{equation}
R(\theta)S(\alpha,\beta)R(\theta)^{-1} = S(\alpha\cos(\theta)+\beta\sin(\theta),
-\alpha\sin(\theta)+\beta\cos(\theta)).
\end{equation}

\begin{theorem}
The states $\Psi_{k,\sigma}$ spanning the subspace acted upon by the
Wigner little group are distinguished by the eigenvalues of $J_{3}=J_{12}$:
\begin{equation}
J_{3}\Psi_{k,\sigma} = \sigma\Psi_{k,\sigma}.
\end{equation}
The $\sigma$ gives the component of angular momentum in the direction of
motion (since $\vec{k}=\|\vec{k}\|\widehat{z}$).
\end{theorem}

\begin{definition}
The eigenvalue $\sigma$ of $J_{3}$ for the state $\Psi_{k,\sigma}$ is
called \define{Helicity}.
\end{definition}

\begin{theorem}
The irreducible representation of an element $W$ of the Wigner little
group for a massless particle satisfies
\begin{equation}
D_{\sigma',\sigma}(W) = \exp(\I\theta\sigma)\,\delta_{\sigma',\sigma}.
\end{equation}
\end{theorem}

\begin{corollary}
The massless single-particle state transforms under an arbitrary Lorentz
transformation as in Eq~\eqref{eq:rqm:poincare-algebra:one-particle-states:action-of-lorentz-transformation-on-state}
like
\begin{equation}
U(\Lambda)\Psi_{p,\sigma} = \sqrt{\frac{(\Lambda p)^{0}}{p^{0}}}\exp(\I\sigma\theta(\Lambda,p))\,\Psi_{\Lambda p,\sigma},
\end{equation}
where $\theta$ is the angle defined by equating $W$ as in Theorem~\ref{thm:rqm:poincare-algebra:massless-case:general-element-of-little-group}
with $W(\Lambda,p) = L(\Lambda p)^{-1}\Lambda L(p)$.
\end{corollary}

\begin{theorem}
Helicity is a Lorentz-invariant quantity for a massless 1-particle system.
\end{theorem}

\begin{remark}
We can think of the absolute value of the helicity $|\sigma|$ as
classifying the ``species'' of particle. For example, $\sigma=\pm1$
describes a massless gauge boson (like the photon), and $\sigma=\pm2$ describes
the graviton.\index{Graviton}
\end{remark}

\N{References}
A detailed discussion of the Poincar\'e and Lorentz groups (and their
Lie algebras) may be found in Chapter 2 of
Weinberg~\cite{Weinberg:1995mt}.
An ``executive summary'' may be found in Srednicki~\cite{Srednicki:2007qs}.
The discussion found in Straumann~\cite{Straumann:2008kq} is also enlightening.
