\subsection{The ``Classical Electron''}
Before continuing on, I will use the phrase ``classical electron'' to
refer to a sphere of radius
\begin{equation}
r_{e} = \frac{1}{4\pi\varepsilon_{0}}\frac{e^{2}}{m_{e}c}\approx2.81794\times10^{-15}~\mathrm{m}.
\end{equation}
It has uniform electric charge density, summing to a total charge equalling
\begin{equation}
q_{e} = \left(\frac{4}{3}\pi r_{e}^{3}\right)\rho_{q} = e\approx-1.602176634\times10^{-19}~\mathrm{C},
\end{equation}
and similarly has a uniform mass density, summing to a total mass of:
\begin{equation}
m_{e} = \left(\frac{4}{3}\pi r_{e}^{3}\right)\rho_{m} \approx 9.1093837015(28)\times10^{-31}~\mathrm{kg}.
\end{equation}
The total angular momentum from the rotation is $S=\hbar/2$, and the
axis of rotation is determined by the velocity vector.


\begin{exercise}[Classical ``electron'' scattering]
\textit{Fluff:} the spin of the electron contributes meaningfully to
scattering experiments. We will see if this is true classically.

\textit{Parameters:}
Consider a fixed point-particle of charge $q\neq0$.
Suppose we have a solid sphere of radius $r_{e}$ initially a distance
$\ell\gg r_{e}$ away from the fixed point-particle; the sphere has
uniform charge density $\rho_{c}$, uniform mass density $\rho_{m}$, and
rotates about its ``forward direction'' (the unit velocity
vector). Suppose its total angular momentum comes from this spin is $S=\hbar/2$.

%% For concrete values, the total charge of the sphere is equal to that of
%% the electron,
%% \begin{subequations}
%% \begin{equation}
%% q_{e} = \left(\frac{4}{3}\pi r_{e}^{3}\right)\rho_{q} = e\approx-1.602176634\times10^{-19}~\mathrm{C},
%% \end{equation}
%% and the total mass of the sphere is that of the electron,
%% \begin{equation}
%% m_{e} = \left(\frac{4}{3}\pi r_{e}^{3}\right)\rho_{m} \approx 9.1093837015(28)\times10^{-31}~\mathrm{kg}.
%% \end{equation}
%% Suppose the radius of the sphere is the classical electron radius,
%% \begin{equation}
%% r_{e} = \frac{1}{4\pi\varepsilon_{0}}\frac{e^{2}}{m_{e}c}\approx2.81794\times10^{-15}~\mathrm{m}.
%% \end{equation}
%% \end{subequations}
We could suppose $\ell=1.00~\mathrm{m}$, for example, which would
intuitively fit inside any laboratory.

\textit{Problem statement:}
If the solid sphere has impact parameter $b$, then what is its
scattering angle $\Theta$? What is its differential cross-sectional area $\D\sigma$?
How do these quantities depend on the spin $S$?
\end{exercise}

\begin{remark}
Recall, the moment of inertia for a sphere would be $I=\frac{2}{5}mr^{2}$
so its angular momentum would be $S=I\omega$.
\end{remark}

\begin{remark}
Since the ``electron'' is spinning and electrically charged, it will
generate a magnetic field. Does our solution reflect this? If we take
$r_{e}\to0$ limit, will our solution recover the familiar scattering of
a charged point-particle? If not, why not?
\end{remark}