\chapter{Quantum Mechanics}

\M We will review the pertinent aspects of \emph{nonrelativistic} quantum
mechanics. This is pedagogical, in the sense we start with ``simple but wrong''
rules, and later on revise the rules as needed.

\section{Basic Rules}\label{sec:qm:basic-rules}

\M We will review the basic rules for describing a quantum system using
a wave function. We will revise the rules later, for greater generality,
but we begin with relevant versions for physical systems.

\N{Position--Momentum Commutator Relation}
We have
\begin{equation}\label{eq:position-momentum-uncertainty-commutator}
[\widehat{x},\widehat{p}] = \I\hbar.
\end{equation}


\subsection{Wave Functions}

\M Following Isham~\cite{Isham:1995lq}, we summarize the rules for
interpreting wave functions.

\N{First rule}
The quantum state of a point particle moving in one-dimension is
represented by a complex-valued wave function $\psi(x)$ which is
normalized to one:
\begin{equation}
\int^{\infty}_{-\infty}|\psi(x)|^{2}\,\D x = 1.
\end{equation}
We will say that ``$\psi$ is normalized to unity'' when this condition holds.

\N{Superposition principle}
We can superimpose any two wave functions $\psi_{1}(x)$ and
$\psi_{2}(x)$ with arbitrary complex coordinates $\alpha_{1}$, $\alpha_{2}$.
That is to say, we get a new wave function
$\alpha_{1}\psi_{1}(x) + \alpha_{2}\psi_{2}(x)$, provided that
$\alpha_{1}$ and $\alpha_{2}$ are chosen to make the new wave function
normalized to unity.

\N{Second rule}
Any physical quantity which we can measure (i.e., any observable) is
represented by a linear differential operator that acts on wave functions
and is self-adjoint.

\begin{remark}[Eigenvalues of self-adjoint operators]
``Self-adjoint'' operators have real eigenvalues. We measure
eigenvalues. If we are measuring a complex-valued quantity, then we
need an anti-self-adjoint operator.
\end{remark}

\begin{remark}[Completeness of eigenfunctions for self-adjoint operators]
The eigenfunctions for a self-adjoint operator are ``complete'' in the
sense that they form a basis. Further, normalized eigenfunctions satisfy
the orthogonality condition
\begin{equation}
\int^{\infty}_{-\infty}f_{m}^{*}(x)f_{n}(x)\,\D x = \delta_{m,n}
\end{equation}
where $\delta_{m,n}=1$ if $m=n$ and $\delta_{m,n}=0$ otherwise.
\end{remark}

\N{Third rule}
The only possible result of measuring an observable $A$ is one of the
eigenvalues of the self-adjoint operator $\widehat{A}$ which represents
it.

\begin{remark}
This rule implies the long-term average value of the results of repeated
measurements of an observable $A$ is
\begin{equation}
\langle A\rangle_{\psi} = \langle\psi\mid A\mid\psi\rangle = \int^{\infty}_{-\infty}\psi^{*}(x)(\widehat{A}\psi)(x)\,\D x.
\end{equation}
\end{remark}

\M
For nondegenerate operators (i.e., any two eigenfunctions of
$\widehat{A}$ with the same eigenvalue are proportional to each other)
and with a discrete set of eigenvalues $a_{1}$, $a_{2}$, \dots
corresponding to eigenfunctions $f_{1}$, $f_{2}$, \dots; if the state is
$\psi(x)$, then the probability that a measurement of $A$ will yield a
particular eigenvalue $a_{n}$ is
\begin{equation}
\Pr(A = a_{n}; \psi) = |\psi_{n}|^{2},
\end{equation}
assuming the wave function is written as a linear combination of the
eigenfunctions,
\begin{equation}
\psi(x) = \sum_{n}\psi_{n}f_{n}(x)
\end{equation}
and the $\psi_{n}\in\CC$ are constants. We also assume $\psi(x)$ is
normalized to unity.

\begin{remark}
As a consequence of this, a wave function being normalized to unity is
the same as the ``sum of all probabilities equal 1''.
\end{remark}

\N{Fourth Rule}
The state function evolved in time according to the time-dependent
Schr\"{o}dinger equation
\begin{equation}
\I\hbar\frac{\partial\psi(x,t)}{\partial t} = \widehat{H}\psi(x,t),
\end{equation}
where the Hamiltonian operator $\widehat{H}$ is obtained from the
classical enerygy expression
\begin{equation}
H = \frac{p^{2}}{2m} + V(x),
\end{equation}
replacing the momentum $p$ and position $x$ by their corresponding
operators $\widehat{p}$, $\widehat{x}$.

\N{Incompleteness of Rules}
As we have presented things, these rules are incomplete since they give
no information on how to construct the operator which represents any
specific observable for a given physical system. In practice, this is
done by the ``substitution rule'' putting hats on things: for any
classical observable $A(x, p)$ the analogous operator is
$A(\widehat{x},\widehat{p})$. 

\begin{danger}
The astute reader will note we never defined ``observable'',
``measurement'', ``physical system'', ``property'', ``state'',
``causality'', or ``determinism''. We also didn't define ``physical quantity''.
This ambiguity is seldom addressed directly, instead most texts shift
the meaning of these words as needed. What is an ``object''? What is a
``property''? Or a ``physical quantity''? Is the significance of these
terms really so obvious? We're confronted by a deceptive question: ``What
is a `thing'?''
\end{danger}

\input{tex/qm/harmonic-oscillator}
\input{tex/qm/experimental-background}
