%%
%% bib.tex
%% 
%% Made by alex
%% Login   <alex@tomato>
%% 
%% Started on  Tue Feb 14 09:22:17 2012 alex
%% Last update Tue Feb 14 09:22:17 2012 alex
%%
\renewcommand{\leftmark}{References}\phantomsection\addcontentsline{toc}{section}{References}
\renewcommand\refname{Books}
\begin{thebibliography}{99}\phantomsection\addcontentsline{toc}{subsection}{\refname}
\bibitem[BauSh]{baumgarte}
  Thomas W.\ Baumgarte, Stuart L.\ Shapiro,\newblock
  \emph{Numerical Relativity: Solving Einstein's Equations on the Computer}.\newblock
  Cambridge University Press (2010).\newblock
  One of the better textbooks on numerical relativity.
\bibitem[Bes]{besse}
  Arthur L.~Besse,\newblock
  \emph{Einstein manifolds}.\newblock
  Springer (2007) pp.~xii+516; \href{http://dx.doi.org/10.1007/978-3-540-74311-8}{\tt doi:10.1007/978-3-540-74311-8}\newblock
  The most authoritative book on Lorentzian geometry (don't let
  the title fool you!).
\bibitem[Bru]{bruhat}
  Yvonne Choquet-Bruhat,\newblock
  \emph{General Relativity and the Einstein Equations}.\newblock
  Oxford University Press, USA (2009).\newblock
  An authoritative comprehensive text on mathematical relativity.
\bibitem[Carlip]{Carlip:1998uc} 
  S.~Carlip,\newblock
  \emph{Quantum gravity in 2+1 dimensions}.\newblock
  Cambridge University Press (1998) 276 pp.\newblock
  Discusses spin connections, among many other topics, quite beautifully.
  %%CITATION = INSPIRE-488048;%%
\bibitem[Carroll]{Carroll:2004st}
  Sean Carroll,\newblock
  \emph{Spacetime and Geometry: An Introduction to General Relativity}.\newblock
  Benjamin Cummings (2003) 513 pp.
  Eprint \arXiv{gr-qc/9712019}
  %%CITATION = INSPIRE-650093;%%
\bibitem[Chr]{christodoulou}
  Demetrios Christodoulou,\newblock
  \emph{Mathematical Problems of General Relativity}, I.\newblock
  European Mathematical Society (2008)
\bibitem[FelCl]{deFelice:1990hu}
  F.\ de Felice, C.J.S.\ Clarke,\newblock
  \emph{Relativity on Curved Manifolds}.\newblock
  Cambridge University Press (1992)\newblock
  This book approaches general relativity with a ``tetrad first''
  philosophy.\newblock 
  \textsc{warning:\quad}\ignorespaces The book uses arcane
  Bourbaki-esque notation for inverses and cotangent spaces.
  %%CITATION = INSPIRE-312813;%%
\bibitem[Har]{Hartle:2003yu}
  James B. Hartle,\newblock
  \emph{Gravity: An Introduction to Einstein's General Relativity}.\newblock
  Benjamin Cummings (2003) 582 pp.
  %%CITATION = INSPIRE-639409;%%
\bibitem[Hat1]{hatcher1}
  Allen Hatcher,\newblock
  \emph{Algebraic Topology}.\newblock
  Eprint: \url{http://www.math.cornell.edu/~hatcher/AT/ATpage.html}
\bibitem[Hat2]{hatcher2}
  \bysame,\newblock
  \emph{Vector Bundles \& K-Theory}.\newblock
  Eprint: \url{http://www.math.cornell.edu/~hatcher/VBKT/VBpage.html}
\bibitem[HawEll]{Hawking:1973uf}
  Stephen W. Hawking, G.F.R.\ Ellis,\newblock
  \emph{The Large Scale Structure of Space-Time}.\newblock
  Cambridge University Press (1975).
  %%CITATION = INSPIRE-87997;%%
\bibitem[Jack]{jackson:1999}
  J.D. Jackson,\newblock
  \emph{Classical electrodynamics}.\newblock
  Wiley; Third Edition (1999).
\bibitem[LPPT]{lightman:1975}
  Alan P. Lightman, William H. Press, Richard H. Price, Saul A. Teukolsky,\newblock
  \emph{Problem Book in Relativity and Gravitation}.\newblock
  Princeton University Press (1975).
\bibitem[MTW]{Misner:1974qy}
  Charles Misner, Kip Thorne, John Wheeler,\newblock
  \emph{Gravitation}.\newblock
  W. H. Freeman; First Edition (1973).\newblock
  General Relativity's Bible.
  %%CITATION = INSPIRE-95654;%%
\bibitem[PR1]{Penrose:1985jw} 
  Roger Penrose and Wolfgang Rindler,\newblock
  \emph{Spinors And Space-time, vol 1: Two Spinor Calculus And Relativistic Fields}.\newblock
  First ed., Cambridge University Press (1984) 458 pp.
  %%CITATION = INSPIRE-216889;%%
\bibitem[PR2]{Penrose:1986ca}
  \bysame, \bysame,\newblock
  \emph{Spinors and Space-Time, vol 2: Spinor And Twistor Methods In Space-time Geometry}.\newblock
  First ed., Cambridge University Press (1986) 501 pp.
  %%CITATION = INSPIRE-238378;%%
\bibitem[Poi]{poisson}
  Eric Poisson,\newblock
  \emph{A Relativist's Toolkit: The Mathematics of Black-Hole Mechanics}.\newblock
  Cambridge University Press (2004), eprint: \href{http://www.physics.uoguelph.ca/poisson/research/agr.pdf}{{\texttt{physics.uoguelph.ca/poisson/research/agr.pdf}}}
\bibitem[Ring]{ringstrom}
  Hans Ringstrom,\newblock
  \emph{The Cauchy Problem in General Relativity}.\newblock
  European Mathematical Society (2009).
\bibitem[Schutz1]{schutz1}
  Bernard Schutz,\newblock
  \emph{Geometrical Methods of Mathematical Physics}.\newblock
  Cambridge University Press; First Published edition (1980).
\bibitem[Schutz2]{Schutz:1985jx}
  \bysame,\newblock
  \emph{A First Course in General Relativity}.\newblock
  Cambridge University Press (1985).
  %%CITATION = INSPIRE-216890;%%
\bibitem[Sharpe]{sharpe}
  R.W.~Sharpe,\newblock
  \emph{Differential Geometry: Cartan's Generalization of Klein's Erlangen Program}.\newblock
  Springer--Verlag (1997).
\bibitem[Spi]{spivak:2010}
  Michael Spivak,\newblock
  \emph{Physics for Mathematicians, Mechanics I}.\newblock
  Publish or Perish; first edition (2010). 749 pp.\newblock
  Great introduction to classical mechanics for mathematicians.
\bibitem[SKMHH]{Stephani:2003tm}
  Hans Stephani, Dietrich Kramer, Malcolm MacCallum, Cornelius Hoenselaers, Eduard Herlt,\newblock
  \emph{Exact Solutions of Einstein's Field Equations}.\newblock
  Cambridge University Press; 2d edition (2009)
  %%CITATION = INSPIRE-619666;%%
\bibitem[Ste]{Stewart:1990uf}
  John Stewart,\newblock
  \emph{Advanced General Relativity}.\newblock
  Cambridge University Press (1990).
  %%CITATION = INSPIRE-312823;%%
\bibitem[Wald]{Wald:1984rg}
  Robert M.\ Wald,\newblock
  \emph{General Relativity}.\newblock
  University Of Chicago Press; First Edition edition (1984)
  %%CITATION = INSPIRE-209356;%%
\bibitem[Wass]{wasserman}
  Robert H.\ Wasserman,\newblock
  \emph{Tensors and Manifolds: With Applications to Physics}.\newblock
  Second ed., Oxford University Press (2009).
\bibitem[Wei]{Weinberg:2008zzc} 
  S.~Weinberg,\newblock
  \emph{Cosmology}.\newblock
  Oxford University Press (2008) 593 pp.
  %%CITATION = INSPIRE-794379;%%
\end{thebibliography}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand\refname{Articles}
\begin{thebibliography}{99}\phantomsection\addcontentsline{toc}{subsection}{\refname}
\bibitem{Anderson:1998we} 
  A.~Anderson and J.~W.~York, Jr.,
  ``Hamiltonian time evolution for general relativity,''\newblock
  \emph{Phys.\ Rev.\ Lett.} {\bf 81} no.~6 (1998) 1154--1157\newblock
  [\arXiv{gr-qc/9807041}].
  %%CITATION = GR-QC/9807041;%%
\bibitem{Aasi:2012fw}
  J.~Aasi {\it et al.}  [The LIGO Scientific and the Virgo Collaboration],\newblock
  ``Einstein@Home all-sky search for periodic gravitational waves in LIGO S5 data,''\newblock
  Eprint \arXiv[gr-qc]{1207.7176}, 29 pages.\newblock
  %%CITATION = ARXIV:1207.7176;%%
\bibitem{Baez:2001qy}
  J.~C.~Baez and Emory F.~Bunn,\newblock
  ``The Meaning of Einstein's Equation,''\newblock
  \emph{Am.\ J.\ Phys.}  {\bf 73} no.~7 (2005) 644--652\newblock
  [\arXiv{gr-qc/0103044}].
  %%CITATION = AJPIA,73,644;%%
\bibitem{Barcelo:2005fc}
  C.~Barcelo, S.~Liberati and M.~Visser,\newblock
  ``Analogue gravity,''\newblock
  \emph{Living Rev.\ Rel.} {\bf 8} (2005) 12\newblock
  [\arXiv{gr-qc/0505065}].
  %%CITATION = GR-QC/0505065;%%
\bibitem{Bartnik:2002cw} 
  R.~Bartnik and J.~Isenberg,\newblock
  ``The Constraint equations,''\newblock
  In \emph{The Einstein Equations and the Large Scale Behavior of Gravitational Fields: 50 years of the Cauchy Problem in General Relativity} (eds.\ Piotr
  Chrusciel and Helmut Friedrich), Springer 2004.\newblock
  Eprint \arXiv{gr-qc/0405092}, 34 pp.
  %%CITATION = GR-QC/0405092;%%
\bibitem{Bartnik:2004ii} 
  R.~Bartnik,\newblock
  ``Phase space for the Einstein equations,''\newblock
  \emph{Commun.Anal.Geom.}~{\bf13} no.~5 (2005) 845--885\newblock
  [\arXiv{gr-qc/0402070}].
  %%CITATION = GR-QC/0402070;%%
\bibitem{Bartnik:2005hh} 
  R.~Bartnik and P.~Tod,
  ``A Note on static metrics,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 23} no.~2 (2006) 569--572\newblock
  [\arXiv{gr-qc/0512097}].
  %%CITATION = GR-QC/0512097;%%
\bibitem{Baumgarte:1997eg} 
  T.~W.~Baumgarte, G.~B.~Cook, M.~A.~Scheel, S.~L.~Shapiro and S.~A.~Teukolsky,\newblock
  ``General relativistic models of binary neutron stars in quasiequilibrium,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 57} no.~12 (1998) 7299--7311\newblock
  [\arXiv{gr-qc/9709026}].
  %%CITATION = GR-QC/9709026;%%
\bibitem{Bertschinger:1994nc}
  E.~Bertschinger and A.~J.~S.~Hamilton,\newblock
  ``Lagrangian evolution of the Weyl tensor,''\newblock
  \emph{Astrophys.\ J.} {\bf 435} no.~1 (1994) 1--7\newblock
  [\arXiv{astro-ph/9403016}].
  %%CITATION = ASTRO-PH/9403016;%%
\bibitem{blandford}
  Roger Blandford and Kip Thorne,\newblock
  \emph{Applications of Classical Physics}.\newblock
  Eprint: \url{http://www.pma.caltech.edu/Courses/ph136/yr2011/}\newblock
  Lecture notes for Caltech physics 136 course series
  ``Applications of Classical Physics''.\newblock
  Great resource showing how fantastically differential geometry
  describes classical physics.
\bibitem{Chamseddine:2005td} 
  A.~H.~Chamseddine,\newblock
  ``Applications of the gauge principle to gravitational interactions,''\newblock
  \emph{Int.\ J.\ Geom.\ Meth.\ Mod.\ Phys.} {\bf 3} (2006) 149--176.\newblock
  [\arXiv{hep-th/0511074}].
  %%CITATION = HEP-TH/0511074;%%
\bibitem{Christodoulou:2008nj} 
  D.~Christodoulou,\newblock
  ``The Formation of Black Holes in General Relativity,''\newblock
  Eprint: \arXiv[gr-qc]{0805.3880}, 594 pp.
  %%CITATION = ARXIV:0805.3880;%%
\bibitem{Chrusciel:1998hc} 
  P.~T.~Chrusciel,\newblock
  ``Mathematical studies of field equations: GR15 Workshop A3,''\newblock
  Eprint \arXiv{gr-qc/9803008}, 25 pages.
  %%CITATION = GR-QC/9803008;%%
\bibitem{Chrusciel:2004cc} 
  P.~T.~Chrusciel, J.~Isenberg and D.~Pollack,\newblock
  ``Initial data engineering,''\newblock
  \emph{Commun.\ Math.\ Phys.} {\bf 257} no.~1 (2005) 29--42\newblock
  [\arXiv{gr-qc/0403066}].
  %%CITATION = GR-QC/0403066;%%
\bibitem{Chrusciel:2010fn} 
  P.~T.~Chrusciel, G.~J.~Galloway and D.~Pollack,\newblock
  ``Mathematical general relativity: a sampler,''\newblock
  \emph{Bull. Amer. Math. Soc.} {\bf 47} (2010), 567--638.\newblock
  \arXiv[gr-qc]{1004.1016} 87 pp.
  %%CITATION = ARXIV:1004.1016;%%
\bibitem{Claret:2010cj} 
  A.~Claret, G.~Torres and M.~Wolf,\newblock
  ``DI Her as a test of internal stellar structure and General Relativity: New apsidal motion rate and evolutionary models,''\newblock
  \emph{Astronomy \& Astrophysics} \textbf{515}, A4 (2010)\newblock
  \arXiv[astro-ph.SR]{1002.2949}.
  %%CITATION = ARXIV:1002.2949;%%
\bibitem{Clifton:2006jh}
  T.~Clifton,\newblock
  ``Alternative theories of gravity,''\newblock
  Doctoral thesis, 205 pages. Eprint \arXiv{gr-qc/0610071}.
  %%CITATION = GR-QC/0610071;%%
%\cite{Eichmair:2012jk}
\bibitem{Eichmair:2012jk} 
  M.~Eichmair, G.~J.~Galloway and D.~Pollack,\newblock
  ``Topological censorship from the initial data point of view,''\newblock
  Eprint \arXiv[gr-qc]{1204.0278}.
  %%CITATION = ARXIV:1204.0278;%%
\bibitem{Elizalde:2004xa}
  E.~Elizalde,\newblock
  ``Cosmology: Techniques and observations,''\newblock
  Eprint \arXiv{gr-qc/0409076}, 64 pages.\newblock
  Lectures to nonspecialists.
  %%CITATION = GR-QC/0409076;%%
\bibitem{Ellis:1994md}
  G.~F.~R.~Ellis and P.~K.~S.~Dunsby,\newblock
  ``Newtonian evolution of the Weyl tensor,''\newblock
  \emph{Astrophys.\ J.} {\bf 479} no.~1 (1997) 97--101\newblock
  [\arXiv{astro-ph/9410001}].
  %%CITATION = ASTRO-PH/9410001;%%
\bibitem{EspositoFarese:1996si}
  G.~Esposito-Farese,\newblock
  ``Binary pulsar tests of strong field gravity,''\newblock
  Eprint \arXiv{gr-qc/9612039}, 18 pages.
  %%CITATION = GR-QC/9612039;%%
\bibitem{Ferrando:2008sr}
  J.~J.~Ferrando and J.~A.~Saez,\newblock
  ``On the algebraic types of the Bel-Robinson tensor,''\newblock
  \emph{Gen.\ Rel.\ Grav.} {\bf 41} no.~8 (2009) 1695--1715\newblock
  [\arXiv[gr-qc]{0807.0181}].
  %%CITATION = ARXIV:0807.0181;%%
\bibitem{Finster:1998ws}
  F.~Finster, J.~Smoller and S.~-T.~Yau,\newblock
  ``Particle - like solutions of the Einstein-Dirac equations,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 59} no.~10 (1999) 104020\newblock
  [\arXiv{gr-qc/9801079}].
  %%CITATION = GR-QC/9801079;%%
\bibitem{Friedrich:2000qv} 
  H.~Friedrich and A.~D.~Rendall,\newblock
  ``The Cauchy problem for the Einstein equations,''\newblock
  \emph{Lect.\ Notes Phys.}  {\bf 540} (2000) pp.~127--224\newblock
  [\arXiv{gr-qc/0002074}].
  %%CITATION = GR-QC/0002074;%%
\bibitem{Garcia:2003bw} 
  A.~Garcia, F.~W.~Hehl, C.~Heinicke and A.~Macias,\newblock
  ``The Cotton tensor in Riemannian space-times,''\newblock
  \emph{Class.\ Quant.\ Grav.}  {\bf 21} no.~4 (2004) 1099--1118\newblock
  [\arXiv{gr-qc/0309008}].
  %%CITATION = GR-QC/0309008;%%
\bibitem{gibbons:2010}
  Gary W.\ Gibbons,\newblock
  ``Special Relativity.''\newblock
  Lecture notes available {\tt\href{http://www.damtp.cam.ac.uk/research/gr/members/gibbons/gwgPartI_SpecialRelativity2010.pdf}{online}}
\bibitem{Giulini:2006uy} 
  D.~Giulini,\newblock
  ``Algebraic and geometric structures of special relativity,''\newblock
  \emph{Lect.\ Notes Phys.} {\bf 702} (2006) 45--111\newblock
  [\arXiv{math-ph/0602018}].
  %%CITATION = MATH-PH/0602018;%%
\bibitem{Giulini:2008pu} 
  \bysame,\newblock%D.~Giulini,\newblock
  ``The Rich Structure of Minkowski Space,''\newblock
  \emph{Fundamental Theories of Physics} {\bf 165}, (2010) 83--132.\newblock
  \arXiv[math-ph]{0802.4345}.
  %%CITATION = ARXIV:0802.4345;%%
\bibitem{gundlach}
  Jens H. Gundlach, Stephen M. Merkowitz, \newblock
  ``Measurement of Newton's Constant Using a Torsion Balance with Angular Acceleration Feedback.''\newblock
  \emph{Phys.\ Rev.\ Lett.}  {\bf 85}, 2869 (2000)\newblock
  [\arXiv{gr-qc/0006043}].
  %%CITATION = GR-QC/0006043;%%
\bibitem{Hsu:2010vp} 
  S.~D.~H.~Hsu,\newblock
  ``White holes and eternal black holes,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 29} no.~1 (2012) 015004\newblock
  [\arXiv[gr-qc]{1007.2934}].
  %%CITATION = ARXIV:1007.2934;%%
\bibitem{Isenberg:2005xp} 
  J.~Isenberg, D.~Maxwell and D.~Pollack,\newblock
  ``A Gluing construction for non-vacuum solutions of the Einstein constraint equations,''\newblock
  \emph{Adv.\ Theor.\ Math.\ Phys.} {\bf 9} no.~1 (2005) 129--172\newblock
  [\arXiv{gr-qc/0501083}].
  %%CITATION = GR-QC/0501083;%%
\bibitem{Kar:2006ms} 
  S.~Kar and S.~SenGupta,\newblock
  ``The Raychaudhuri equations: A Brief review,''\newblock
  \emph{Pramana} {\bf 69} no.~1 (2007) 49--76\newblock
  [\arXiv{gr-qc/0611123}].
  %%CITATION = GR-QC/0611123;%%
\bibitem{Kramer:2004gj} 
  M.~Kramer, 2, A.~G.~Lyne, M.~Burgay, A.~Possenti, R.~N.~Manchester, F.~Camilo, M.~A.~McLaughlin and D.~R.~Lorimer {\it et al.},\newblock
  ``The Double pulsar. A New testbed for relativistic gravity,''\newblock
  In \emph{Binary Pulsars} (Eds.\ Rasio \& Stairs), PASP.\newblock
  [\arXiv{astro-ph/0405179}].
  %%CITATION = ASTRO-PH/0405179;%%
\bibitem{Laarakkers:1997hb} 
  W.~G.~Laarakkers and E.~Poisson,\newblock
  ``Quadrupole moments of rotating neutron stars,''\newblock
  \emph{Astrophys.~J.}~{\bf512} no.~1 (1999) 282--287.\newblock
  [\arXiv{gr-qc/9709033}].
  %%CITATION = GR-QC/9709033;%%
\bibitem{Martin:2012bt}
  J.~Martin,\newblock
  ``Everything You Always Wanted To Know About The Cosmological Constant Problem (But Were Afraid To Ask),''\newblock
  Eprint \arXiv[astro-ph.CO]{1205.3365}, 89 pages.
  %%CITATION = ARXIV:1205.3365;%%
\bibitem{CODATA2002}
  Peter J. Mohr; Barry N. Taylor, \newblock
  ``CODATA recommended values of the fundamental physical constants: 2002''\newblock
  \emph{Reviews of Modern Physics} {\bf77} no.~1 (2005) 1--107.\newblock
  Section Q (pp.\ 42--47) describes the mutually inconsistent measurement experiments from which the CODATA value for $G$ was derived.
  %%CITATION = INSPIRE-603889;%%
\bibitem{Morsink:2007tv} 
  S.~M.~Morsink, D.~A.~Leahy, C.~Cadeau and J.~Braga,\newblock
  ``The Oblate Schwarzschild Approximation for Light Curves of Rapidly Rotating Neutron Stars,''\newblock
  \emph{Astrophys.\ J.}  {\bf 663} no.~2 (2007) 1244--1251\newblock
  [\arXiv{astro-ph/0703123}].
  %%CITATION = ASTRO-PH/0703123;%%
\bibitem{nelson:rudimentary}
  Alex Nelson,\newblock
  ``Rudimentary notes on Differential Geometry.''\newblock
  Eprint \url{http://code.google.com/p/notebk/files/rudimentaryDifferentialGeometry.v0.pdf},
  i+16 pages.
\bibitem{Nieto:2007ng}
  M.~M.~Nieto,\newblock
  ``The Quest to Understand the Pioneer Anomaly,''\newblock
  Eprint \arXiv{gr-qc/0702017}, 12 pages.
  %%CITATION = GR-QC/0702017;%%
\bibitem{Peldan:1993hi} 
  P.~Peldan,\newblock
  ``Actions for gravity, with generalizations: A Review,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 11} no.~5 (1994) 1087--1132\newblock
  [\arXiv{gr-qc/9305011}].
  %%CITATION = GR-QC/9305011;%%
\bibitem{Pons:2009cz} 
  J.~M.~Pons, D.~C.~Salisbury and K.~A.~Sundermeyer,\newblock
  ``Revisiting observables in generally covariant theories in the light of gauge fixing methods,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 80} no.~8 (2009) 084015\newblock
  [\arXiv[gr-qc]{0905.4564}].
  %%CITATION = ARXIV:0905.4564;%%
\bibitem{Pons:2010ad} 
  \bysame, \bysame and \bysame,\newblock
  ``Observables in classical canonical gravity: folklore demystified,''\newblock
  \emph{J.\ Phys.\ Conf.\ Ser.} {\bf 222} no.~1 (2010) 012018\newblock
  [\arXiv[gr-qc]{1001.2726}].
  %%CITATION = ARXIV:1001.2726;%%
\bibitem{Rovelli:2001my} 
  C.~Rovelli,\newblock
  ``GPS observables in general relativity,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 65} no.~4 (2002) 044017\newblock
  [\arXiv{gr-qc/0110003}].
  %%CITATION = GR-QC/0110003;%%
\bibitem{Steinhoff:2010zz} 
  J.~Steinhoff,\newblock
  ``Canonical formulation of spin in general relativity,''\newblock
  \emph{Annalen Phys.} {\bf 523} no.~4 (2011) 296--353\newblock
  [\arXiv[gr-qc]{1106.4203}].
  %%CITATION = ARXIV:1106.4203;%%
\bibitem{Straumann:2005hj} 
  N.~Straumann,\newblock
  ``Gauge principle and QED,''\newblock
  \emph{Acta Phys.\ Polon.\ B} {\bf 37} no.~3 (2006) 575--594\newblock
  [\arXiv{hep-ph/0509116}].
  %%CITATION = HEP-PH/0509116;%%
\bibitem{Tambornino:2011vg} 
  J.~Tambornino,\newblock
  ``Relational Observables in Gravity: a Review,''\newblock
  \emph{Symmetry, Integrability and Geometry: Methods and Applications} \textbf{8} (2012) 017\newblock
  Eprint \arXiv[gr-qc]{1109.0740}.
  %%CITATION = ARXIV:1109.0740;%%
\bibitem{Torre:1993fq} 
  C.~G.~Torre,\newblock
  ``Gravitational observables and local symmetries,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 48} no.~6 (1993) 2373--2376\newblock
  [\arXiv{gr-qc/9306030}].
  %%CITATION = GR-QC/9306030;%%
\bibitem{Townsend:1997ku} 
  P.~K.~Townsend,\newblock
  ``Black holes: Lecture notes,''\newblock
  [\arXiv{gr-qc/9707012}]. 145 pp.
  %%CITATION = GR-QC/9707012;%%
\bibitem{Vallisneri:2008ye} 
  M.~Vallisneri,\newblock
  ``A LISA Data-Analysis Primer,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 26} no.~9 (2009) 094024\newblock
  Eprint \arXiv[gr-qc]{0812.0751} 13 pages.
  %%CITATION = ARXIV:0812.0751;%%
\bibitem{Will:1998dx} 
  Clifford M. Will,\newblock
  ``The Confrontation between General Relativity and Experiment: A 1998 Update.''\newblock
  \emph{eConf C} {\bf 9808031}, 02 (1998)\newblock
  Eprint: \arXiv{gr-qc/9811036}, 76 pages.
  %%CITATION = GR-QC/9811036;%%
\bibitem{Williams:2005rv} 
  James G. Williams, Slava G. Turyshev, Dale H. Boggs,\newblock
  ``Lunar Laser Ranging Tests of the Equivalence Principle with the Earth and Moon.''\newblock
  \emph{Int.\ J.\ Mod.\ Phys.\ D} {\bf 18} no.~7 (2009) 1129--1175\newblock
  Eprint: \arXiv{gr-qc/0507083}, 50 pages.
  %%CITATION = GR-QC/0507083;%%
\bibitem{Winn:2005ef} 
  J.~N.~Winn, R.~W.~Noyes, M.~J.~Holman, D.~Charbonneau, Y.~Ohta, A.~Taruya, Y.~Suto and N.~Narita {\it et al.},\newblock
  ``Measurement of spin-orbit alignment in an extrasolar planetary system,''\newblock
  \emph{Astrophys.\ J.} {\bf 631} no.~2 (2005) 1215--1226\newblock
  [\arXiv{astro-ph/0504555}].
  %%CITATION = ASTRO-PH/0504555;%%
\bibitem{Woodard:2006nt} 
  R.~P.~Woodard,\newblock
  ``Avoiding dark energy with $1/r$ modifications of gravity,''\newblock
  \emph{Lect.\ Notes Phys.} {\bf 720} (2007) 403--433\newblock
  [\arXiv{astro-ph/0601672}].
  %%CITATION = ASTRO-PH/0601672;%%
\end{thebibliography}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand\refname{Advanced Relativity}
\begin{thebibliography}{99}\phantomsection\addcontentsline{toc}{subsection}{\refname}
\bibitem{Alekseev:2010mx}
  G.~A.~Alekseev,\newblock
  ``Thirty years of studies of integrable reductions of Einstein's field equations,''\newblock
  Eprint \arXiv[gr-qc]{1011.3846}, 22 pages.
  %%CITATION = ARXIV:1011.3846;%%
\bibitem{Bernal:2003jb}
  A.~N.~Bernal and M.~Sanchez,\newblock
  ``On Smooth Cauchy hypersurfaces and Geroch's splitting theorem,''\newblock
  \emph{Commun.\ Math.\ Phys.} {\bf 243} no.~3 (2003) 461--470\newblock
  [\arXiv{gr-qc/0306108}].
  %%CITATION = GR-QC/0306108;%%
\bibitem{Bicak:1997bx}
  J.~Bicak and K.~V.~Kuchar,\newblock
  ``Null dust in canonical gravity,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 56} (1997) 4878--4895\newblock
  [\arXiv{gr-qc/9704053}].
  %%CITATION = GR-QC/9704053;%%
\bibitem{Bicak:2000ea}
  J.~Bicak,\newblock
  ``Selected solutions of Einstein's field equations: Their role in general relativity and astrophysics,''\newblock
  \emph{Lect.\ Notes Phys.} {\bf 540} (2000) 1--126\newblock
  [\arXiv{gr-qc/0004016}].
  %%CITATION = GR-QC/0004016;%%
%\cite{Brown:1994su}
\bibitem{Blohmann:2010jd}
  C.~Blohmann, M.~C.~B.~Fernandes and A.~Weinstein,\newblock
  ``Groupoid symmetry and constraints in general relativity,''\newblock
  Eprint \arXiv[math.DG]{1003.2857}, 22 pages.
  %%CITATION = ARXIV:1003.2857;%%

\bibitem{Brown:2000dz}
  J.~D.~Brown, S.~R.~Lau and J.~W.~York, Jr.,\newblock
  ``Action and energy of the gravitational field,''\newblock
  Eprint \arXiv{gr-qc/0010024}, 50 pages.
  %%CITATION = GR-QC/0010024;%%
\bibitem{Cardoso:2010ac}
  J.~G.~Cardoso,\newblock
  ``The Classical Two-Component Spinor Formalisms for General Relativity,''\newblock
  Eprint \arXiv[math-ph]{1004.5150}, 51 pages.
  %%CITATION = ARXIV:1004.5150;%%
\bibitem{Castagnino:2012iz}
  M.~Castagnino and L.~Chimento,\newblock
  ``Two Theorems on Flat Space-Time Gravitational Theories,''\newblock
  \emph{Gen.\ Rel.\ Grav.} {\bf 12} (1980) 825--835\newblock
  [\arXiv[gr-qc]{1206.5341}].
  %%CITATION = ARXIV:1206.5341;%%
\bibitem{ChoquetBruhat:2012jz}
  Y.~Choquet-Bruhat and P.~T.~Chrusciel,\newblock
  ``Cauchy problem with data on a characteristic cone for the Einstein-Vlasov equations,''\newblock
  Eprint \arXiv[gr-qc]{1206.0390}, 14 pages.
  %%CITATION = ARXIV:1206.0390;%%
\bibitem{Clayton:1998ai}
  M.~A.~Clayton,\newblock
  ``Canonical general relativity: Matter fields in a general linear frame,''\newblock
  \emph{J.\ Math.\ Phys.} {\bf 40} (1999) 3476\newblock
  [\arXiv{gr-qc/9808005}].
  %%CITATION = GR-QC/9808005;%%
\bibitem{Damour:2007uh}
  T.~Damour,\newblock
  ``General Relativity Today,''\newblock
  Eprint \arXiv[gr-qc]{0704.0754}, 49 pages.
  %%CITATION = ARXIV:0704.0754;%%
\bibitem{Deser:1976ay}
  S.~Deser and C.~J.~Isham,\newblock
  ``Canonical Vierbein Form of General Relativity,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 14} (1976) 2505.\newblock
  Although this paper doesn't have a free eprint, its historical
  importance must be stressed!
  %%CITATION = PHRVA,D14,2505;%%
\bibitem{Franke:2007ux}
  V.~A.~Franke,\newblock
  ``Different canonical formulations of Einstein's theory of gravity,''\newblock
  \emph{Theor.\ Math.\ Phys.} {\bf 148} no.~1 (2006) 995--1010\newblock
  [\arXiv[gr-qc]{0710.4953}].
  %%CITATION = ARXIV:0710.4953;%%
\bibitem{Friedrich:2002xz}
  H.~Friedrich,\newblock
  ``Conformal Einstein evolution,''\newblock
  \emph{Lect.\ Notes Phys.} {\bf 604} (2002) 1--50\newblock
  [\arXiv{gr-qc/0209018}].
  %%CITATION = GR-QC/0209018;%%
\bibitem{Frolov:2008sn}
  A.~M.~Frolov, N.~Kiriushcheva and S.~V.~Kuzmin,\newblock
  ``On canonical transformations between equivalent Hamiltonian formulations of General Relativity,''\newblock
  \emph{Grav.\ Cosmol.} {\bf 17} (2011) 314\newblock
  [\arXiv[gr-qc]{0809.1198}].
  %%CITATION = ARXIV:0809.1198;%%
\bibitem{Gielen:2010cu}
  S.~Gielen and D.~Oriti,\newblock
  ``Classical general relativity as BF-Plebanski theory with linear constraints,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 27} (2010) 185017\newblock
  [\arXiv[gr-qc]{1004.5371}].
  %%CITATION = ARXIV:1004.5371;%%
\bibitem{Gielen:2010nb}
  S.~Gielen,\newblock
  ``Classical GR as a topological theory with linear constraints,''\newblock
  \emph{J.\ Phys.\ Conf.\ Ser.} {\bf 314} (2011) 012044\newblock
  [\arXiv[gr-qc]{1011.5513}].
  %%CITATION = ARXIV:1011.5513;%%
\bibitem{Jaramillo:2010ay}
  J.~L.~Jaramillo and E.~Gourgoulhon,\newblock
  ``Mass and Angular Momentum in General Relativity,''\newblock
  \emph{Fundam.\ Theor.\ Phys.} {\bf 162} (2011) 87
  [\arXiv[gr-qc]{1001.5429}].
  %%CITATION = ARXIV:1001.5429;%%
\bibitem{Kiriushcheva:2008fn}
  N.~Kiriushcheva, S.~V.~Kuzmin, C.~Racknor and S.~R.~Valluri,\newblock
  ``Diffeomorphism Invariance in the Hamiltonian formulation of General Relativity,''\newblock
  \emph{Phys.\ Lett.\ A} {\bf 372} (2008) 5101\newblock
  [\arXiv[gr-qc]{0808.2623}].
  %%CITATION = ARXIV:0808.2623;%%
\bibitem{Kiriushcheva:2008sf}
  N.~Kiriushcheva and S.~V.~Kuzmin,\newblock
  ``The Hamiltonian formulation of General Relativity: Myths and reality,''\newblock
  \emph{Central Eur.\ J.\ Phys.} {\bf 9} (2011) 576--615\newblock
  [\arXiv[gr-qc]{0809.0097}].
  %%CITATION = ARXIV:0809.0097;%%
\bibitem{Kiriushcheva:2011ct}
  N.~Kiriushcheva, P.~G.~Komorowski and S.~V.~Kuzmin,\newblock
  ``Lagrangian symmetries of the ADM action. Do we need a solution to the `non-canonicity puzzle'?''\newblock
  Eprint \arXiv[gr-qc]{1108.6105}.
  %%CITATION = ARXIV:1108.6105;%%
\bibitem{Kluson:2012tw}
  J.~Kluson,\newblock
  ``Hamiltonian Analysis of the Conformal Decomposition of the Gravitational Field,''\newblock
  Eprint \arXiv[gr-qc]{1206.5116}, 12 pages.
  %%CITATION = ARXIV:1206.5116;%%
\bibitem{Muller:2009bw}
  T.~Muller and F.~Grave,\newblock
  ``Catalogue of Spacetimes,''\newblock
  Eprint \arXiv[gr-qc]{0904.4184}, 88 pages.
  %%CITATION = ARXIV:0904.4184;%%
\bibitem{Pervushin:2012gz}
  V.~N.~Pervushin, A.~B.~Arbuzov, B.~M.~Barbashov, R.~G.~Nazmitdinov, A.~Borowiec, K.~N.~Pichugin and A.~F.~Zakharov,\newblock
  ``Conformal and Affine Hamiltonian Dynamics of General Relativity,''\newblock
  \emph{ General Relativity and Gravitation}
  Eprint \arXiv[gr-qc]{1109.2789} .
  %%CITATION = ARXIV:1109.2789;%%
\bibitem{Poplawski:2009fb}
  N.~J.~Poplawski,\newblock
  ``Spacetime and Fields,''\newblock
  Eprint \arXiv[gr-qc]{0911.0334}, 114 pages.
  %%CITATION = ARXIV:0911.0334;%%
%\cite{Rendall:2002dr}
\bibitem{Rendall:1997dc}
  A.~D.~Rendall,\newblock
  ``Global dynamics of the Mixmaster model,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 14} no.~8 (1997) 2341--2356\newblock
  [\arXiv{gr-qc/9703036}].
  %%CITATION = GR-QC/9703036;%%
\bibitem{Rendall:2002dr}
  \bysame, \newblock%A.~D.~Rendall,\newblock
  ``Theorems on existence and global dynamics for the Einstein equations,''\newblock
  \emph{Living Rev.\ Rel.} {\bf 5} (2002) 6\newblock
  [\arXiv{gr-qc/0203012}].
  %%CITATION = GR-QC/0203012;%%
\bibitem{Rendall:2005nf}
  \bysame,\newblock%A.~D.~Rendall,\newblock
  ``The Nature of spacetime singularities,''\newblock
  Eprint \arXiv{gr-qc/0503112}, 19 pages.
  %%CITATION = GR-QC/0503112;%%
\bibitem{Reula:2010yt}
  O.~Reula and O.~Sarbach,\newblock
  ``The Initial-Boundary Value Problem in General Relativity,''\newblock
  \emph{Int.\ J.\ Mod.\ Phys.\ D} {\bf 20} no.~5 (2011) 767--783\newblock
  [\arXiv[gr-qc]{1009.0589}].
  %%CITATION = ARXIV:1009.0589;%%
\bibitem{Steinhoff:2008zr}
  J.~Steinhoff, G.~Schaefer and S.~Hergt,\newblock
  ``ADM canonical formalism for gravitating spinning objects,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 77} (2008) 104018\newblock
  [\arXiv[gr-qc]{0805.3136}].
  %%CITATION = ARXIV:0805.3136;%%
\bibitem{Steinhoff:2010zz}
  J.~Steinhoff,\newblock
  ``Canonical formulation of spin in general relativity,''\newblock
  \emph{Annalen Phys.} {\bf 523} (2011) 296--353\newblock
  [\arXiv[gr-qc]{1106.4203}].
  %%CITATION = ARXIV:1106.4203;%%
\bibitem{Straumann:2000ke}
  N.~Straumann,\newblock
  ``Reflections on gravity,''\newblock
  Eprint \arXiv{astro-ph/0006423}, 26 pages.\newblock
  Presents classical General Relativity as generated with
  classical spin-2 particle on an unobservable flat spacetime.  
  %%CITATION = ASTRO-PH/0006423;%%
\end{thebibliography}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand\refname{Quantum Gravity}
\begin{thebibliography}{99}\phantomsection\addcontentsline{toc}{subsection}{\refname}
% Overviews
\bibitem{Isham:1993ji}
  C.~J.~Isham,\newblock
  ``Prima facie questions in quantum gravity,''\newblock
  In \emph{Bad Honnef 1993, Proceedings, Canonical gravity} 1--21; and London Imp. Coll. - ICTP-93-94-01 (93/10,rec.Nov.) 21 p\newblock
  [\arXiv{gr-qc/9310031}].
  %%CITATION = GR-QC/9310031;%%
\bibitem{DeWitt:2007mi}
  B.~S.~DeWitt and G.~Esposito,\newblock
  ``An Introduction to quantum gravity,''\newblock
  \emph{Int.\ J.\ Geom.\ Meth.\ Mod.\ Phys.} {\bf 5} no.~1 (2008) 101--156\newblock
  [\arXiv[hep-th]{0711.2445}]
  %%CITATION = ARXIV:0711.2445;%%
\bibitem{Esposito:2011rx}
  G.~Esposito,\newblock
  ``An Introduction to quantum gravity,''\newblock
  Section 6.7.17 of the EOLSS Encyclopedia by UNESCO\newblock
  [\arXiv[hep-th]{1108.3269}]
  %%CITATION = ARXIV:1108.3269;%%

% Phenomenological aspects
\bibitem{Bonder:2012gf}
  Y.~Bonder,\newblock
  ``Lorentz Invariant Phenomenological Model of Quantum Gravity: A Minimalistic Presentation,''\newblock
  Eprint \arXiv[gr-qc]{1204.0055}, 5 pages.
  %%CITATION = ARXIV:1204.0055;%%
\bibitem{Marciante:2012iu}
  M.~Marciante and T.~Schucker,\newblock
  ``Fluctuation of Dirac operator and equivalence principle,''\newblock
  Eprint \arXiv[hep-th]{1203.4960}, 49 pages.
  %%CITATION = ARXIV:1203.4960;%%
\bibitem{Wolf:2010na}
  P.~Wolf, L.~Blanchet, C.~J.~Borde, S.~Reynaud, C.~Salomon and C.~Cohen-Tannoudji,\newblock
  ``Does an atom interferometer test the gravitational redshift at the Compton frequency?''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 28} no.~14 (2011) 145017\newblock
  [\arXiv[gr-qc]{1012.1194}].
  %%CITATION = ARXIV:1012.1194;%%
\bibitem{Wolf:2011im}
  \bysame, \bysame, \bysame, \bysame, \bysame{} and \bysame,\newblock
  ``Testing the Gravitational Redshift with Atomic Gravimeters?''\newblock
  Eprint \arXiv[gr-qc]{1106.3412}, 5 pages.
  %%CITATION = ARXIV:1106.3412;%%

% semiclassical gravity
\bibitem{Giulini:2011uw}
  D.~Giulini and A.~Grossardt,\newblock
  ``Gravitationally induced inhibitions of dispersion according to the Schr\"odinger-Newton Equation,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 28} no.~19 (2011) 195026\newblock
  [\arXiv[gr-qc]{1105.1921}].
  %%CITATION = ARXIV:1105.1921;%%
\bibitem{Giulini:2012zu}
  D.~Giulini and A.~Grosardt,\newblock
  ``The Schr\"odinger-Newton equation as non-relativistic limit of self-gravitating Klein-Gordon and Dirac fields,''\newblock
  \arXiv[gr-qc]{1206.4250}.
  %%CITATION = ARXIV:1206.4250;%%
\bibitem{Hack:2012qf}
  T.P.~Hack and V.~Moretti,\newblock
  ``On the Stress-Energy Tensor of Quantum Fields in Curved Spacetimes --- Comparison of Different Regularization Schemes and Symmetry of the Hadamard/Seeley-DeWitt Coefficients,''\newblock
  Eprint \arXiv[gr-qc]{1202.5107}, 24 pages.
  %%CITATION = ARXIV:1202.5107;%%
\bibitem{Date:2011rd}
  G.~Date,\newblock
  ``Revisiting canonical gravity with fermions,''\newblock
  Eprint \arXiv[gr-qc]{1110.3416}, 29 pages.
  %%CITATION = ARXIV:1110.3416;%%
\bibitem{Kiefer:1993fg}
  C.~Kiefer,\newblock
  ``The Semiclassical approximation to quantum gravity,''\newblock
  In \emph{Bad Honnef 1993, Proceedings, Canonical gravity} 170--212, and Freiburg U. - THEP-93-27 (93/12,rec.Dec.) 48 pp\newblock
  [\arXiv{gr-qc/9312015}].
  %%CITATION = GR-QC/9312015;%%

% QFT in Curved Spacetime
\bibitem{GarciaCompean:2011zt}
  H.~Garcia-Compean and F.~J.~Turrubiates,\newblock
  ``Ground-state Wigner functional of linearized gravitational field,''\newblock
  \emph{Int.\ J.\ Mod.\ Phys.} {\bf A26} no.~30 (2011) 5241--5259\newblock
  [\arXiv[hep-th]{1109.1036}].
  %%CITATION = ARXIV:1109.1036;%%



% Canonical approaches to Quantum Gravity
\bibitem{Hamber:2012zm}
  H.~W.~Hamber, R.~Toriumi and R.~M.~Williams,\newblock
  ``Wheeler-DeWitt Equation in 2 + 1 Dimensions,''\newblock
  Eprint \arXiv[hep-th]{1207.3759}, 56 pages.
  %%CITATION = ARXIV:1207.3759;%%
\bibitem{Giulini:1994dx}
  D.~Giulini and C.~Kiefer,\newblock
  ``Wheeler-DeWitt metric and the attractivity of gravity,''\newblock
  \emph{Phys.\ Lett.\ A} {\bf 193} (1994) 21--24\newblock
  [\arXiv{gr-qc/9405040}].
  %%CITATION = GR-QC/9405040;%%
\bibitem{Gielen:2011qq}
  S.~Gielen,\newblock
  ``The Space of Connections as the Arena for (Quantum) Gravity,''\newblock
  \emph{SIGMA} {\bf 7} (2011) 104\newblock
  [\arXiv[gr-qc]{1111.2672}].
  %%CITATION = ARXIV:1111.2672;%%
\bibitem{Isham:1992ms}
  C.~J.~Isham,\newblock
  ``Canonical quantum gravity and the problem of time,''\newblock
  In \emph{Salamanca 1992, Proceedings, Integrable systems, quantum groups, and quantum field theories} 157--287, and London Imp. Coll. - ICTP-91-92-25 (92/08,rec.Nov.) 124 p\newblock
  [\arXiv{gr-qc/9210011}].
  %%CITATION = GR-QC/9210011;%%
\bibitem{Kuchar:1993ne}
  K.~V.~Kuchar,\newblock
  ``Canonical quantum gravity,''\newblock
  Eprint \arXiv{gr-qc/9304012}, 35 pages.
  %%CITATION = GR-QC/9304012;%%





% Relating Canonical approaches to others
\bibitem{Aros:2003bi}
  R.~Aros, M.~Contreras and J.~Zanelli,\newblock
  ``Path integral measure for first order and metric gravities,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 20} no.~13 (2003) 2937--2944\newblock
  [\arXiv{gr-qc/0303113}].
  %%CITATION = GR-QC/0303113;%%
\bibitem{Barvinsky:1997hp}
  A.~O.~Barvinsky and C.~Kiefer,\newblock
  ``Wheeler-DeWitt equation and Feynman diagrams,''\newblock
  \emph{Nucl.\ Phys.\ B} {\bf 526} (1998) 509--539\newblock
  [\arXiv{gr-qc/9711037}].
  %%CITATION = GR-QC/9711037;%%
\bibitem{Vaz:2002xb}
  C.~Vaz, C.~Kiefer, T.~P.~Singh and L.~Witten,\newblock
  ``Quantum general relativity and Hawking radiation,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 67} (2003) 024014\newblock
  [\arXiv{gr-qc/0208083}].
  %%CITATION = GR-QC/0208083;%%

% Path Integrals 
\bibitem{Alexandrov:1998cu}
  S.~Y.~.Alexandrov and D.~V.~Vassilevich,\newblock
  ``Path integral for the Hilbert-Palatini and Ashtekar gravity,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 58} (1998) 124029\newblock
  [\arXiv{gr-qc/9806001}].
  %%CITATION = GR-QC/9806001;%%
\bibitem{Barnich:1996sh}
  G.~Barnich and V.~Husain,\newblock
  ``Geometrical representation of the constraints of Euclidean general relativity,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 14} no.~5 (1997) 1043--1058\newblock
  [\arXiv{gr-qc/9611030}].
  %%CITATION = GR-QC/9611030;%%
\bibitem{Barvinsky:1997fk}
  A.~O.~Barvinsky,\newblock
  ``Solution of quantum Dirac constraints via path integral,''\newblock
  \emph{Nucl.\ Phys.\ B} {\bf 520} (1998) 533--560\newblock
  [\arXiv{hep-th/9711164}].
  %%CITATION = HEP-TH/9711164;%%
\bibitem{Kshirsagar:1992fc}
  A.~K.~Kshirsagar,\newblock
  ``Towards a path integral for pure spin connection formulation of gravity,''\newblock
  \emph{Class.\ Quant.\ Grav.} {\bf 10} no.~9 (1993) 1859--1864\newblock
  [\arXiv{hep-th/9207115}].
  %%CITATION = HEP-TH/9207115;%%
\bibitem{MenaMarugan:1993eq}
  G.~A.~Mena Marugan,\newblock
  ``Reality conditions for Lorentzian and Euclidean gravity in the Ashtekar formulation,''\newblock
  \emph{Int.\ J.\ Mod.\ Phys.\ D} {\bf 3} (1994) 513--528\newblock
  [\arXiv{gr-qc/9311020}].
  %%CITATION = GR-QC/9311020;%%


% Quantum Black Hole Stuff
\bibitem{Brown:1994su}
  J.~D.~Brown and J.~W.~York, Jr.,\newblock
  ``The Path integral formulation of gravitational thermodynamics,''\newblock
  In \emph{The Black Hole: 25 Years After} (eds.~Teitelboim, \emph{et al.}) pp.~1--24\newblock
  [\arXiv{gr-qc/9405024}].\newblock
  Discusses the Euclidean action as the entropy.
  %%CITATION = GR-QC/9405024;%%

% Quantum Cosmology
\bibitem{Husain:2003ry}
  V.~Husain and O.~Winkler,\newblock
  ``On singularity resolution in quantum gravity,''\newblock
  \emph{Phys.\ Rev.\ D} {\bf 69} (2004) 084016\newblock
  [\arXiv{gr-qc/0312094}]. Obtains results similar to Loop
  Quantum Gravity's big bounce model.
  %%CITATION = GR-QC/0312094;%%
\bibitem{Rinaldi:2012dh}
  M.~Rinaldi,\newblock
  ``Aspects of Quantum Gravity in Cosmology,''\newblock
  \emph{Mod.\ Phys.\ Lett.\ A} {\bf 27} no.~7 (2012) 1230008\newblock
  [\arXiv[gr-qc]{1201.4543}].
  %%CITATION = ARXIV:1201.4543;%%




\bibitem{Halliwell:2012nj}
  J.~J.~Halliwell and J.~M.~Yearsley,\newblock
  ``Pitfalls of Path Integrals: Amplitudes for Spacetime Regions and the Quantum Zeno Effect,''\newblock
  Eprint \arXiv[gr-qc]{1205.3773}.
  %%CITATION = ARXIV:1205.3773;%%

% Historical papers
\bibitem{Giddings:2011xs}
  S.~B.~Giddings,\newblock
  ``The gravitational S-matrix: Erice lectures,''\newblock
  Eprint \arXiv[hep-th]{1105.2036}, 44 pages.
  %%CITATION = ARXIV:1105.2036;%%
\bibitem{Nirov:1994qk}
  K.~.S.~Nirov,\newblock
  ``Constraint algebras in gauge invariant systems,''\newblock
  \emph{Int.\ J.\ Mod.\ Phys.\ A} {\bf 10} (1995) 4087\newblock
  [\arXiv{hep-th/9407156}].
  %%CITATION = HEP-TH/9407156;%%
\end{thebibliography}
