%%
%% bib.tex
%% 
%% Made by alex
%% Login   <alex@tomato>
%% 
%% Started on  Tue Dec 27 11:25:08 2011 alex
%% Last update Tue Dec 27 11:25:08 2011 alex
%%

\begin{thebibliography}{99}\addcontentsline{toc}{section}{References}
\bibitem{graphs}
J.L. Gross and T.W. Tucker, \newblock
\emph{Topological graph theory}. \newblock
Wiley Interscience, 1987
\bibitem{bourbaki} Nicolas Bourbaki,
\newblock \emph{General Topology: Chapters 1--4}.
\newblock Springer (1998) 452 pages.
\newblock Useful reference for topological groups.
\bibitem{hatcher2006algebraic} Allen Hatcher,
\newblock\emph{Algebraic Topology}.
\newblock Cambridge University Press (2002).
\newblock Eprint: \url{http://www.math.cornell.edu/~hatcher/AT/ATpage.html}
xii+550 pages.
\bibitem{may} J. P. May, \newblock
\emph{A Concise Course in Algebraic Topology}.\newblock
Eprint: \url{http://www.math.uchicago.edu/~may/CONCISE/ConciseRevised.pdf}
\bibitem{milnor} John Milnor,\newblock
``On Spaces Having the Homotopy Type of a CW-Complex''.\newblock
\emph{Transactions of the AMS} {\bf 90} 2 (1959) 272--280.
\bibitem{munkres} J.R. Munkres. \newblock
\emph{Elements of algebraic topology}.\newblock
Addison--Wesley (1984).
\bibitem{nelson} Alex Nelson, \newblock
``A Rapid Introduction to Topology.''\newblock
Eprint: \url{http://notebk.googlecode.com/files/rapidIntroTopology.pdf},
6 pages.
\bibitem{nelson2} \bysame, \newblock
``Lie Groups Notes.''\newblock
Eprint: \url{http://notebk.googlecode.com/files/lieGroups.lieAlgebras.Representations.gamma.pdf},
91 pages.
\bibitem{schwarz} Albert Schwarz,
\newblock\emph{Topology for Physicists}.
\newblock Springer--Verlag (2010) 307 pages. 
\bibitem{steenrod} Norman Steenrod,
\newblock \emph{The Topology of Fibre Bundles}
\newblock Princeton Landmarks in Mathematics \textbf{14}.
\newblock Princeton University Press (1999) 224 pages.
\end{thebibliography}
